---
title: "BCRW_simulations"
author: "Scott Forrest"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages 

```{r packages} 
# , message=F, warning=F

library(tidyverse)
packages <- c("amt", "lubridate", "terra", "tictoc", "RColorBrewer", "patchwork", 
              "sf", "ggpubr","Rfast", "beepr", "SpatialKDE", "ks", "circular")
walk(packages, require, character.only = T)

```


Import buffalo telemetry data.

```{r message = FALSE}

# buffalo_CLR <- read_csv(file = "outputs/buffalo_parametric_indv_covs_20230208.csv")
# using the population estimate as the TENTATIVE distribution will make it easier to estimate the weighted mean estimate of the movement kernel
buffalo_data <- read_csv(file = "outputs/buffalo_parametric_popn_covs_20230208.csv")

unique(buffalo_data$id)

# scaling_mean <- mean(values(ndvi_projected[[7]], na.rm = TRUE)) # 0.4054526 for July 2018
# scaling_sd <- sd(values(ndvi_projected[[7]], na.rm = TRUE)) # 0.1113644 for July 2018

buffalo_CLR <- buffalo_data %>% 
  mutate(id_num = as.numeric(factor(id)), 
         step_id = step_id_, 
         x1 = x1_, x2 = x2_, 
         y1 = y1_, y2 = y2_, 
         t1 = t1_, t1_rounded = round_date(buffalo_data$t1_, "hour"), t2 = t2_, 
         sl = sl_, log_sl = log_sl_, 
         ta = ta_, cos_ta = cos_ta_,
         year = year(t1_), 
         month = month(t1_), 
         # month_factor = factor(month(t1_)),
         hour = hour(t1_),
         # hour_factor = factor(hour),
         # day = day(t1_),
         # day_factor = factor(day(t1_)),
         # step_id_factor = factor(step_id_),
         ndvi_scaled = ((ndvi_temporal - 0.4054526) / 0.1113644),
         canopy_01 = canopy_cover/100,
         herby_scaled = scale(veg_herby),
         canopy_scaled = scale(canopy_01),
         elev_scaled = scale(DEM_H_end),
         elev_delta_scaled = scale(elev_delta),
         elev_log_scaled = scale(elev_log),
         slope_scaled = scale(slope_end),
         month_s1 = sin(2*pi*month/12),
         month_s2 = sin(4*pi*month/12),
         month_c1 = cos(2*pi*month/12),
         month_c2 = cos(4*pi*month/12)) %>%
  drop_na(c(ndvi_temporal, veg_herby, canopy_01, sl_)) %>% 
  dplyr::select(!(burst_:case_))

```

Determining how many individuals have at least 1 year of data

```{r}

# buffalo_CLR %>% ggplot(aes(x = t1, y = factor(id), colour = factor(id))) +
#   geom_point(alpha = 0.1) +
#   scale_y_discrete("Buffalo ID") +
#   scale_x_datetime("Date") +
#   scale_colour_viridis_d() +
#   theme_bw() +
#   theme(legend.position = "none")
# 
# buffalo_CLR %>% dplyr::group_by(id) %>%  
#   summarise(min_time = min(t1), max_time = max(t1),
#             min_x = min(x2), max_x = max(x2),
#             min_y = min(y2), max_y = max(y2))

```


Filtering by animals that have more than a year of data

```{r}

# these individuals have more than 1 year of data (or very close to)
buffalo_year_ids <- c(2005, 2014, 2018, 2022, 2024, 2154, 2158, 2327, 2354, 2387)
buffalo_CLR_year <- buffalo_CLR %>% filter(id %in% buffalo_year_ids)
buffalo_CLR_year_pres <- buffalo_CLR_year %>% filter(y == 1) %>% dplyr::select(c(id, y, x1, x1, y1, y2, t1, month, hour))

# write_csv(buffalo_CLR_year %>% dplyr::select(c(id, y, x1, x1, y1, y2, t1, month)), "outputs/buffalo_CLR_year_locations_20230220.csv")

buffalo_CLR_year <- read_csv("outputs/buffalo_CLR_year_locations_20230220.csv")

```


Setting up before the simulation
- resource layers
- coefficient values
- random starting locations

```{r}

# monthly_coefs <- read_csv("outputs/buffalo_parametric_indv_movement_params_20230208.csv")
monthly_coefs <- read_csv("outputs/monthly_CLR_coefs_wide_with_movement_params_20230210.csv") # this is for the weighted mean
monthly_coefs$date <- with_tz(monthly_coefs$date, tzone = "Australia/Queensland")

```


Resources

```{r}

ndvi_above0_stack_scaled <- rast("mapping/cropped rasters/ndvi_above0_stack_scaled_20230210.tif")
slope_scaled <- rast("mapping/cropped rasters/slope_scaled_by_buffalo_data_20230210.tif")
veg_herby <- rast("mapping/cropped rasters/veg_herby.tif")
canopy_scaled <- rast("mapping/cropped rasters/canopy_scaled_by_buffalo_data_20230210.tif")

```


Generate habitat surfaces to sample from (i.e. the naive approach)

```{r temporal naive}

# create date vector
years <- c(rep(2019, 7), rep(2018, 5))
months <- c(seq(1, 7, 1), seq(8, 12, 1))
day <- rep(1, 12)
dates <- make_datetime(year = years, month = months, day = day, tz = "Australia/Queensland")

# set up loop
layers_naive <- c()

for(i in 1:12) {
  
  # i = 1

resources <- c(ndvi_above0_stack_scaled[[which(time(ndvi_above0_stack_scaled) == monthly_coefs$date[i])]], 
               slope_scaled,
               veg_herby, 
               canopy_scaled)

# plot(resources)
# extent(resources) <- c(0, 60000, 0, 57000)

naive <- rast(resources)[[1]]
naive <- terra::setValues(naive, 0)
names(naive) <- "naive_predictions"
  
for (j in 1:4) {
  naive <- naive + monthly_coefs[i,][[j + 1]] * resources[[j]]
}
  
naive_norm <- exp(naive)/global(exp(naive), fun = "sum", na.rm = TRUE)[[1]]
# terra::plot(naive_norm)
layers_naive <- c(layers_naive, naive_norm)

}

naive_predictions <- rast(layers_naive)
plot(naive_predictions[[1]])
points(buffalo_CLR_year_pres$x1, buffalo_CLR_year_pres$y1)

# layers <- mask(layers, water, maskvalue = 1)
# raster::plot(layers, col = brewer.pal(9, "Reds"), main = dates)

terra::time(naive_predictions) <- dates
terra::time(naive_predictions) 

# for(i in 1:12) (plot(naive_predictions[[i]], main = time(naive_predictions[[i]])))

# writeRaster(naive_predictions, "mapping/pred_outputs/naive/naive_20230223/naive_stack_20230223.tif")
naive_predictions <- rast("mapping/pred_outputs/naive/naive_20230223/naive_stack_20230223.tif")

```


```{r}

n_indvs <- 1000 # number of individual 'animals'
n_steps <- 100 # number of steps in the trajectory
n_ch <- 100 # number of steps to be selected from

month_n <- 1

```


Select the appropriate probability surface for the month of interest, and possibly crop to a smaller extent.

```{r}

# resources <- c(ndvi_above0_stack_scaled[[which(time(ndvi_above0_stack_scaled) == monthly_coefs$date[month_n])]],
#                slope_scaled,
#                veg_herby, 
#                canopy_scaled)

# ext(resources) <- c(0, 60000, 0, 57000)
# plot(resources[[1]])
habitat_selection <- naive_predictions[[which(time(naive_predictions) == monthly_coefs$date[month_n])]]
plot(habitat_selection)
# nrow(resources) * ncol(resources)
ext(habitat_selection) # 0, 60000, -1463000, -1406000

# smaller 100 x 100 extent
xmin <- 30000
xmax <- 32500
ymin <- -1436000
ymax <- -1433500

# larger extent
xmin <- 12500
xmax <- 45000
ymin <- -1445000
ymax <- -1427500

crop_extent <- ext(xmin, xmax, ymin, ymax)

resources_cropped <- terra::crop(habitat_selection, crop_extent)
ext(resources_cropped) <- c(xmin - xmin, 
                            xmax - xmin, 
                            ymin - ymin, 
                            ymax - ymin)

terra::plot(resources_cropped)
terra::plot(resources_cropped)
points(buffalo_CLR_year$x1 - xmin, buffalo_CLR_year$y1 - ymin)
nrow(resources_cropped) * ncol(resources_cropped)

# resources_cropped_raster <- raster(resources_cropped)

```

Cropping all layers

```{r}

# naive_predictions
# 
# habitat_selection <- naive_predictions[[which(time(naive_predictions) == monthly_coefs$date[month_n])]]
# plot(habitat_selection)
# # nrow(resources) * ncol(resources)

naive_predictions_cropped <- terra::crop(naive_predictions, crop_extent)

ext(naive_predictions_cropped) <- c(xmin - xmin, 
                                    xmax - xmin, 
                                    ymin - ymin, 
                                    ymax - ymin)

terra::plot(naive_predictions_cropped[[1]])
points(buffalo_CLR_year_pres$x1, buffalo_CLR_year_pres$y1)
# writeRaster(naive_predictions_cropped, "outputs/simulated trajectories/prediction_maps/naive_predictions_cropped_20230222.tif")

resources_cropped <- naive_predictions_cropped[[month_n]]

# naive_raster_cropped <- raster::stack(naive_predictions_cropped)
# plot(naive_raster_cropped)

```



# Global parameters for the simulation model

```{r}

# selecting coef values for that month
coef_values <- monthly_coefs[month_n,][2:5]
coef_values

```


Setup starting locations

```{r}

start_locs <- buffalo_CLR_year %>% filter(y == 1 & month == 1 & id == 2154) %>% mutate(xadj = x1 - xmin, yadj = y1 - ymin) %>% slice(1) %>% select(c(xadj, yadj)) %>% data.frame()

start_init <- cbind(rep(start_locs[[1]], n_indvs), rep(start_locs[[2]], n_indvs))

start_unif <- cbind(runif(n_indvs, min = ext(resources_cropped)[1], max = ext(resources_cropped)[2]), 
                    runif(n_indvs, min = ext(resources_cropped)[3], max = ext(resources_cropped)[4]))

```


BCRW function - parametric step lengths

```{r}

# for testing function - comment out otherwise
# coef <- coef_values
# xy0 <- start_unif
# resc <- resources_cropped

simulate_ssf <- function(n_steps, n_ch, coef, xy0, resc) { 
  
  sl <- rgamma(n_steps * n_ch, shape = monthly_coefs$shape[month_n], scale = monthly_coefs$scale[month_n])
  ta <- as.numeric(Rfast::rvonmises(n_steps * n_ch, m = pi, k = monthly_coefs$kappa[month_n]) - pi)
  # hist(sl)
  # hist(ta)
  steps <- rep(1:n_steps, each = n_ch)
  x_0 <- xy0[1]
  y_0 <- xy0[2]
  x_s <- sl * sin(ta)
  y_s <- sl * cos(ta)
  x <- rep(NA, n_steps)
  y <- rep(NA, n_steps)
  x[1] <- x_0
  y[1] <- y_0
  # multiply resources with selection coef
  # for (i in 1:length(coef)) resc[[i]] <- resc[[i]] * coef[i][[1]]
  
  # for testing function - comment out otherwise
  # plot(resc)
  # i = 2
  
  for (i in 2:n_steps) {
    # adding the modulo operator %% ensures a wrapped/toroid landscape, but the origin must be at (0,0)
    x_pos <- (x[i - 1] + sl[steps == i] * sin(ta[steps == i])) %% ext(resc)[2] 
    y_pos <- (y[i - 1] + sl[steps == i] * cos(ta[steps == i])) %% ext(resc)[4]
    # plot(resc)
    # plot(resources[[1]])
    # points(x = x_pos, y = y_pos)
    
    # sampling environment
    # p <- exp(rowSums(terra::extract(resc, cbind(x_pos, y_pos))))
    
    p <- terra::extract(resc, cbind(x_pos, y_pos))[,1]
    
    p[is.na(p)] <- 1e-10
    w <- sample(n_ch, 1, prob = p)
    # w <- which.max(p) # causes it to get stuck in local maxima
    x[i] <- x_pos[w]
    y[i] <- y_pos[w]
  }
  
  
  # step_extract <- function(i, x, y, sl, ta, steps, resc) {
  # 
  #   x_pos <- (x[i - 1] + sl[steps == i] * sin(ta[steps == i])) %% ext(resc)[2]
  #   y_pos <- (y[i - 1] + sl[steps == i] * cos(ta[steps == i])) %% ext(resc)[4]
  # 
  #   p <- terra::extract(resc, cbind(x_pos, y_pos))[,1]
  # 
  #   p[is.na(p)] <- 1e-10
  #   w <- sample(n_ch, 1, prob = p)
  #   # w <- which.max(p) # causes it to get stuck in local maxima
  #   x[i] <- x_pos[w]
  #   y[i] <- y_pos[w]
  #   
  #   return(x)
  # 
  # }
  # 
  #  step_extract(i = 2:5, x = x, y = y, sl = sl, ta = ta, steps = steps, resc = resc)
  # 
  # map_dbl(2:n_steps, step_extract(i = , x = x, y = y, sl = sl, ta = ta, steps = steps, resc = resc))
  # 
  # 
  # # attempting to map rather than loop
  # map_dbl(2:n_steps, function(i, x = x, y = y, sl = sl, ta = ta, steps = steps, resc = resc) {
  # 
  #   x_pos <- (x[i - 1] + sl[steps == i] * sin(ta[steps == i])) %% ext(resc)[2]
  #   y_pos <- (y[i - 1] + sl[steps == i] * cos(ta[steps == i])) %% ext(resc)[4]
  # 
  #   p <- terra::extract(resc, cbind(x_pos, y_pos))[,1]
  # 
  #   p[is.na(p)] <- 1e-10
  #   w <- sample(n_ch, 1, prob = p)
  #   # w <- which.max(p) # causes it to get stuck in local maxima
  #   x[i] <- x_pos[w]
  #   y[i] <- y_pos[w]
  # 
  # })
  
  
  data_frame(x = x, y = y)
}

```


Running the function

```{r}

tic()
stps <- map(1:n_indvs, function(i)
  
  simulate_ssf(n_steps = n_steps, 
               n_ch = n_ch, 
               coef = coef_values, 
               xy0 = start_unif[i,], 
               resc = resources_cropped)
  )
toc()

beep(sound = 2)

```


Near term predictions (the same starting location)

```{r}

tic()
stps <- map(1:n_indvs, function(i)
  
  simulate_ssf(n_steps = n_steps, 
               n_ch = n_ch, 
               coef = coef_values, 
               xy0 = start_init[i,], 
               resc = resources_cropped)
  )
toc()

beep(sound = 2)

```



Generating a dataframe from the list elements (of each individual's trajectory)

```{r}

animals <- data_frame(
  id = paste0("a", 1:n_indvs),
  track = map(stps, ~ track(x = .$x, y = .$y, t = ymd_hm("2019-01-01 00:00") + hours(1:nrow(.))))
)

sim_data <- unnest(animals, cols = track)

write_csv(sim_data, paste("outputs/simulated trajectories/bcrw_buffalo_w_mean_croppedresc_month", month_n,
                          "_", n_indvs, "indvs_", n_ch, "ch_", n_steps, "stps_", Sys.Date(), ".csv", sep = ""))

```


# Looping over months

```{r}

# crop_extent <- ext(00000, 50000, -1450000, -1410000)

sim_data_list <- vector(mode = "list", length = 12)

tic()
for(j in 1:12) {
  
  month_n <- j
  resources <- naive_predictions_cropped[[j]]

### setting up starting locations
start_unif <- cbind(runif(n_indvs, min = ext(resources)[1], max = ext(resources)[2]), 
                    runif(n_indvs, min = ext(resources)[3], max = ext(resources)[4]))

### running the function
stps <- map(1:n_indvs, function(i)
  
  simulate_ssf(n_steps = n_steps, 
               n_ch = n_ch, 
               coef = monthly_coefs[i,][2:5], 
               xy0 = start_unif[i,], 
               resc = resources)
  )

### turning into track object
animals <- data_frame(
  id = paste0("a", 1:n_indvs),
  track = map(stps, ~ track(x = .$x, y = .$y, t = ymd_hm("2019-01-01 00:00") + hours(1:nrow(.))))
)

### expanding into data frame and saving to csv
sim_data_list[[j]] <- unnest(animals, cols = track)
write_csv(sim_data_list[[j]], paste("outputs/simulated trajectories/bcrw_buffalo_w_mean_loop_month", month_n,
                          "_", n_indvs, "indvs_", n_ch, "ch_", n_steps, "stps_", Sys.Date(), ".csv", sep = ""))

}
toc()

```


# Looping over the same month but saving intermittently

```{r}

ext(naive_predictions_cropped) <- c(xmin - xmin, 
                                    xmax - xmin, 
                                    ymin - ymin, 
                                    ymax - ymin)

nrow(naive_predictions_cropped) * ncol(naive_predictions_cropped)

n_indvs <- 100
n_steps <- 1e3
n_ch <- 100
month_n <- 3

sim_data_list <- vector(mode = "list", length = 100)

for(j in 1:100) {
  
  # month_n <- j
  resources <- naive_predictions_cropped[[month_n]] # always the same month

### setting up starting locations
start_unif <- cbind(runif(n_indvs, min = ext(resources)[1], max = ext(resources)[2]), 
                    runif(n_indvs, min = ext(resources)[3], max = ext(resources)[4]))

### running the function
tic()
stps <- map(1:n_indvs, function(i)
  
  simulate_ssf(n_steps = n_steps, 
               n_ch = n_ch, 
               coef = monthly_coefs[month_n,][2:5], 
               xy0 = start_unif[i,], 
               resc = resources)
  )
toc()

### turning into track object
animals <- data_frame(
  id = paste0("a", 1:n_indvs),
  track = map(stps, ~ track(x = .$x, y = .$y, t = ymd_hm("2019-01-01 00:00") + hours(1:nrow(.))))
)

### expanding into data frame and saving to csv
sim_data_list[[j]] <- unnest(animals, cols = track)
write_csv(sim_data_list[[j]], paste("outputs/simulated trajectories/bcrw_buffalo_w_mean_lge_ext_month_run",j , "month_3", month_n,
                          "_", n_indvs, "indvs_", n_ch, "ch_", n_steps, "stps_", Sys.Date(), ".csv", sep = ""))

}

```


Plotting the simulated data

```{r}

sim_data <- sim_data_list[[1]]

coef_values
plot(resources)
plot(resources[[1]])
plot(resources_cropped)
plot(resources_cropped)
points(x = sim_data$x_, y = sim_data$y_)

naive_cropped_df <- as.data.frame(resources_cropped, xy = TRUE)

ggplot() +
  geom_raster(data = naive_cropped_df, aes(x = x, y = y, fill = naive_predictions)) +
  #geom_path(data = sim_data, aes(x = x_, y = y_, colour = id), size = 1) +
  geom_point(data = sim_data, aes(x = x_, y = y_), size = 0.1, colour = "black", alpha = 0.5) + # colour = id
  # geom_point(data = buffalo_subset %>% filter(y == 1), aes(x = x1_, y = y1_ + 1463000), size = 0.1, colour = "red") +
  scale_fill_viridis_c("Habitat suitability (naive)") +
  scale_colour_discrete(guide = "none") +
  #scale_fill_gradient(low = "white", high = "darkgreen") +
  coord_equal() +
  ggtitle("10 steps ahead (n = 1000)") +
  theme_bw() +
  theme(legend.position = "right", legend.key.width = unit(1.1, "cm"))

# ggsave("outputs/plots/bcrw_simulations/bcrw_short_term_preds_20230301.png",
# width=150, height=90, units="mm", dpi = 300)
# ggsave("outputs/plots/bcrw_simulations/bcrw_short_term_preds_pres_20230301.png",
# width=300, height=180, units="mm", dpi = 300, scale = 1)


bcrw_extent <- sim_data %>% filter(id == "a1") %>% 
  summarise(min_x = min(x_), max_x = max(x_),
            min_y = min(y_), max_y = max(y_),
            x_range = max(x_) - min(x_),
            y_range = max(y_) - min(y_))

ggplot() +
  geom_raster(data = naive_cropped_df, aes(x = x, y = y, fill = NDVI)) +
  geom_path(data = sim_data, aes(x = x_, y = y_, colour = id), linewidth = 0.5) +
  geom_point(data = sim_data, aes(x = x_, y = y_, colour = id), size = 0.5, alpha = 0.1) + # colour = "black", 
  # geom_point(data = buffalo_subset %>% filter(y == 1), aes(x = x1_, y = y1_ + 1463000), size = 0.1, colour = "red") +
  scale_fill_viridis_c() +
  scale_colour_discrete(guide = FALSE) +
  #scale_fill_gradient(low = "white", high = "darkgreen") +
  scale_x_continuous(limits = c(bcrw_extent$min_x - 1000, bcrw_extent$max_x + 1000)) +
  scale_y_continuous(limits = c(bcrw_extent$min_y - 1000, bcrw_extent$max_y + 1000)) +
  coord_equal() +
  ggtitle("Simulation Approach") +
  theme_bw() +
  theme(legend.position = "bottom", legend.key.width = unit(1.1, "cm"))

```


## generating distribution map

```{r}

points <- resources_cropped #resource[[1]]
points[] <- 0

sim_ud_points <- as.matrix(sim_data[,2:3])
counts <- as.data.frame(table(terra::cellFromXY(resources_cropped, sim_ud_points))) # resources[[1]]
points[as.numeric(levels(counts$Var1))] <- counts$Freq

plot(points)
# plot(points)
# points(x = sim_data$x, y = sim_data$y)  
plot(resources_cropped)



ggplot() +
  geom_raster(data = naive_df, aes(x = x, y = y, fill = naive_predictions)) +
  #geom_path(data = sim_data, aes(x = x_, y = y_, colour = id), size = 1) +
  # geom_point(data = sim_data, aes(x = x_, y = y_), size = 1, colour = "black", alpha = 0.05) + # colour = id
  # geom_point(data = buffalo_subset %>% filter(y == 1), aes(x = x1_, y = y1_ + 1463000), size = 0.1, colour = "red") +
  scale_fill_viridis_c() +
  # scale_colour_discrete(guide = FALSE) +
  #scale_fill_gradient(low = "white", high = "darkgreen") +
  coord_equal() +
  ggtitle("Naive Approach") +
  theme_bw()

```


# Kernel density estimation

```{r}

sim_m1_100indvs_10ch_10000stps <- read_csv("outputs/simulated trajectories/bcrw_buffalo_w_mean_100indvs_10000stps_2023-02-10.csv")
plot(sim_m1_100indvs_10ch_10000stps$x_, sim_m1_100indvs_10ch_10000stps$y_)

sim_data <- sim_m1_100indvs_10ch_10000stps

sim_data %>% filter(id == unique(sim_data$id)[1:10])

sim_data_sf <- sim_data %>% # %>% filter(id == unique(sim_data$id)[1:10]) 
  st_as_sf(coords = c("x_", "y_"), dim = "XY") %>%
  st_set_crs(3112) %>%
  select()

cell_size <- 25
band_width <- 50

sim_data_raster <- sim_data_sf %>%
  create_raster(cell_size = cell_size, side_offset = band_width)

tic()
sim_data_kde <- sim_data_sf %>%
  SpatialKDE::kde(band_width = band_width, kernel = "triweight", grid = sim_data_raster)
toc()

beep(sound = 2)

plot(rast(sim_data_kde))
sim_data_kde_trim <- crop(rast(sim_data_kde), ext(resources_cropped))
plot(sim_data_kde_trim)

# normalise to sum to 1
sim_data_kde_trim_norm <- sim_data_kde_trim / global(sim_data_kde_trim, fun = "sum", na.rm = TRUE)[[1]]
global(sim_data_kde_trim_norm, fun = "sum", na.rm = TRUE)

resources_cropped_norm <- resources_cropped / global(resources_cropped, fun = "sum", na.rm = TRUE)[[1]]
global(resources_cropped_norm, fun = "sum", na.rm = TRUE)

points_norm <- points / global(points, fun = "sum", na.rm = TRUE)[[1]]
global(points_norm, fun = "sum", na.rm = TRUE)[[1]]

plot(resources_cropped_norm)
plot(points_norm)
plot(sim_data_kde_trim_norm)

```


```{r}

# ext(resources_cropped_norm) <- ext(30000, 32000, -1436000, -1434000)
# ext(points_norm) <- ext(30000, 32000, -1436000, -1434000)
# ext(sim_data_kde_trim_norm) <- ext(30000, 32000, -1436000, -1434000)

# naive_cropped_df <- as.data.frame(resources_cropped_norm, xy = TRUE)
# naive_predictions_cropped_df <- as.data.frame(naive_predictions_cropped, xy = TRUE)
# month(with_tz(time(naive_predictions_cropped), tzone = "Australia/Queensland"))
ext(resources_cropped) <- c(xmin - xmin, xmax - xmin, ymin - ymin, ymax - ymin)
naive_cropped_df <- as.data.frame(resources_cropped, xy = TRUE)


# points_df <- as.data.frame(points_norm, xy = TRUE)
# sim_data_kde_df <- as.data.frame(sim_data_kde_trim_norm, xy = TRUE)

naive_cropped_df
points_df
sim_data_kde_df

### NAIVE ###

ggplot() +
  geom_raster(data = naive_cropped_df, aes(x = x, y = y, fill = naive_predictions)) +
  geom_point(data = buffalo_CLR_year %>% filter(y == 1 & month == 1 & id == 2154), 
             aes(x = x1 - xmin, y = y1 - ymin, colour = factor(id))) +
             # ,
             # colour = "red") +
  coord_equal() +
  scale_x_continuous(limits = c(0, xmax - xmin)) +
  scale_y_continuous(limits = c(0, ymax - ymin)) +
  scale_fill_viridis_c("Naive predictions", option = "D") +
  theme_classic() +
  theme(legend.position = "right")

ggsave("outputs/plots/bcrw_simulations/naive_cropped_wpoints_20230217png", width=150, height=90, units="mm", dpi = 300)

for(i in 1:12) print(ggplot() +
  geom_raster(data = as.data.frame(naive_predictions_cropped_df[,c(1,2,2+i)]), 
              aes(x = x, y = y, fill = !!sym(colnames(as.data.frame(naive_predictions_cropped_df[2+i]))))) +
    scale_fill_viridis_c(option = "D") +
  geom_point(data = buffalo_CLR_year %>% filter(y == 1 & month == i), aes(x = x1, y = y1, colour = factor(id))) +
  coord_equal() +
  scale_x_continuous(limits = c(20000, 40000)) +
  scale_y_continuous(limits = c(-1440000, -1430000)) +
  theme_classic())


### RAW BCRW POINTS ###

ggplot() +
  geom_raster(data = points_df, aes(x = x, y = y, fill = naive_predictions)) +
  geom_point(data = buffalo_CLR_year %>% filter(y == 1 & month == 1), aes(x = x1, y = y1)) +
  coord_equal() +
  scale_x_continuous(limits = c(30000, 32000)) +
  scale_y_continuous(limits = c(-1436000, -1434000)) +
  scale_fill_viridis_c(option = "D") +
  theme_classic()


### KDE BCRW POINTS ###

ggplot() +
  geom_raster(data = sim_data_kde_df, aes(x = x, y = y, fill = layer)) +
  geom_path(data = buffalo_CLR_year %>% filter(y == 1 & month == 1), aes(x = x1, y = y1, colour = factor(id))) +
  coord_equal() +
  scale_x_continuous(limits = c(30000, 32000)) +
  scale_y_continuous(limits = c(-1436000, -1434000)) +
  scale_fill_viridis_c(option = "D") +
  theme_classic()

for(i in 1:12) print(ggplot() +
  geom_raster(data = sim_data_kde_df, aes(x = x, y = y, fill = layer)) +
  geom_path(data = buffalo_CLR_year %>% filter(y == 1 & month == i), aes(x = x1, y = y1, colour = factor(id))) +
  coord_equal() +
  scale_x_continuous(limits = c(30000, 32000)) +
  scale_y_continuous(limits = c(-1436000, -1434000)) +
  scale_fill_viridis_c(option = "D") +
  theme_classic())

```



## Determining an appropriate extent

```{r}

naive_df <- as.data.frame(naive_predictions[[1]], xy = TRUE)
min(naive_df$y, na.rm = TRUE)

ggplot() +
  geom_raster(data = naive_df, 
              aes(x = x, y = y + 1462963, fill = naive_df[,3])) +
  scale_fill_viridis_c("Habitat suitability", direction = 1, 
                       option = "D") + 
    geom_point(data = buffalo_CLR_year %>% filter(month == 1 & y == 1), 
               aes(x = x1, y = y1 + 1462963, colour = factor(id)), size = 1, alpha = 0.75) + # colour = "red", 
  coord_equal() +
  scale_x_continuous(limits = c(30000, 35000)) +
  scale_y_continuous(limits = c(25000, 30000)) +
  # ggtitle(paste("Naive approach", time(naive_predictions_01[[i]]), sep = " ")) +
  theme_bw() +
  theme(legend.position = "right", legend.key.width = unit(1, "cm"))

ggplot()

```







```{r}

points_ud <- points / cellStats(points, stat = "sum")
cellStats(points_ud, stat = "sum")
points_df <- as.data.frame(points_ud, xy = TRUE)
points_df$layer <- points_df$ndvi

bcrw <- 
  ggplot() +
  geom_raster(data = points_df, aes(x = x, y = y, fill = layer)) +
  #geom_point(data = animals_ssf %>% filter(case_ == TRUE), aes(x = x1_, y = y1_, colour = id), size = 0.5) +
  #scale_fill_viridis_c() +
  scale_fill_gradient(low = "white", high = "red") +
  coord_equal() +
  ggtitle("Simulation Approach") +
  theme_bw() +
  theme(legend.position = "bottom", legend.key.width = unit(1.1, "cm"))

bcrw

```









# Currently unused

BCRW function - empirical movement kernel (bootstrapping data)

```{r}

# empirical_steps <- buffalo_all_covs %>% filter(y == 1) %>% .$sl_ 
# empirical_angles <- buffalo_all_covs %>% filter(y == 1) %>% .$ta_
# empirical_pairs <- bind_cols("steps" = empirical_steps, "angles" = empirical_angles) %>% drop_na()

temporal_movement_params$shape[1]

# slice_sample(empirical_pairs, n = 5, replace = T)

n_steps <- 10
n_ch <- 100
coef <- cov_values
xy0 <- start_unif
resc <- resources

simulate_ssf <- function(n_steps, n_ch, coef, xy0, resc) { 
  #sl <- rexp(n_steps * n_ch, rate = l) # l = 1, add into arguments if using exponential distribution for sl
  sl <- rgamma(n_steps * n_ch, shape = temporal_movement_params$shape[1], scale = temporal_movement_params$scale[1])
  ta <- rvonmises(n_steps * n_ch, m = 0, k = temporal_movement_params$kappa[1]) - pi
  # hist(ta)
  #ta <- runif(n_steps * n_ch, -pi, pi)
  # sl <- unlist(as.vector(slice_sample(empirical_pairs, n = n_steps * n_ch, replace = T)[,1]))
  # ta <- unlist(as.vector(slice_sample(empirical_pairs, n = n_steps * n_ch, replace = T)[,2]))
  #stp_ta <- slice_sample(empirical_pairs, n = n_steps * n_ch, replace = T)
  #sl <- as.vector(stp_ta[,1])
  #ta <- as.numeric(stp_ta[,2])
  #sl <- base::sample(x = empirical_steps, size = n_steps * n_ch, replace = TRUE)
  #ta <- base::sample(x = empirical_angles, size = n_steps * n_ch, replace = TRUE)
  steps <- rep(1:n_steps, each = n_ch)
  x_0 <- xy0[1]
  y_0 <- xy0[2]
  x_s <- sl * sin(ta)
  y_s <- sl * cos(ta)
  x <- rep(NA, n_steps)
  y <- rep(NA, n_steps)
  x[1] <- x_0
  y[1] <- y_0
  # multiply resources with selection coef
  for (i in 1:length(coef)) resc[[i]] <- resc[[i]] * coef[i][[1]]
  
  # i = 2
  
  for (i in 2:n_steps) {
    # adding the modulo operator %% ensures a wrapped/toroid landscape, but the origin must be at (0,0)
    x_pos <- (x[i - 1] + sl[steps == i] * sin(ta[steps == i])) %% ext(resc)[2] 
    y_pos <- (y[i - 1] + sl[steps == i] * cos(ta[steps == i])) %% ext(resc)[4]
    p <- exp(rowSums(terra::extract(resc, cbind(x_pos, y_pos))))
    p[is.na(p)] <- 1e-6
    w <- sample(n_ch, 1, prob = p)
    x[i] <- x_pos[w]
    y[i] <- y_pos[w]
  }
  data_frame(x = x, y = y)
}

```