---
title: "SSA CLR sw_window 230324"
author: "Scott Forrest"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import packages

```{r}
#message = FALSE

options(scipen=999)

library(tidyverse)

packages <- c("amt", "lubridate", "mgcv", "survival", "terra", "raster", "tictoc", 
              "RColorBrewer", "patchwork", "ecospat", "ggmap", "basemaps", "mapedit", 
              "sf", "ggpubr", "circular", "glmmTMB")

walk(packages, require, character.only = T)

```
Import buffalo telemetry data

```{r message = FALSE}

# buffalo_CLR <- read_csv(file = "outputs/buffalo_parametric_indv_covs_20230208.csv")
# using the population estimate as the TENTATIVE distribution will make it easier to estimate the weighted mean estimate of the movement kernel
buffalo_data <- read_csv(file = "outputs/buffalo_parametric_popn_covs_20230208.csv")

unique(buffalo_data$id)

# scaling_mean <- mean(values(ndvi_projected[[7]], na.rm = TRUE)) # 0.4054526 for July 2018
# scaling_sd <- sd(values(ndvi_projected[[7]], na.rm = TRUE)) # 0.1113644 for July 2018

buffalo_CLR <- buffalo_data %>% 
  mutate(id_num = as.numeric(factor(id)), 
         step_id = step_id_, 
         x1 = x1_, x2 = x2_, 
         y1 = y1_, y2 = y2_, 
         t1 = t1_, 
         t1_rounded = round_date(buffalo_data$t1_, "hour"), 
         t2 = t2_, 
         yday = yday(t1_),
         sl = sl_, log_sl = log_sl_, 
         ta = ta_, cos_ta = cos_ta_,
         year = year(t1_), 
         month = month(t1_), 
         # month_factor = factor(month(t1_)),
         hour = hour(t1_),
         # hour_factor = factor(hour),
         # day = day(t1_),
         # day_factor = factor(day(t1_)),
         # step_id_factor = factor(step_id_),
         ndvi_scaled = scale(ndvi_temporal),
         ndvi_scaled_raster = ((ndvi_temporal - 0.4054526) / 0.1113644),
         canopy_01 = canopy_cover/100,
         herby_scaled = scale(veg_herby),
         canopy_scaled = scale(canopy_01),
         elev_scaled = scale(DEM_H_end),
         elev_delta_scaled = scale(elev_delta),
         elev_log_scaled = scale(elev_log),
         slope_scaled = scale(slope_end),
         month_s1 = sin(2*pi*month/12),
         month_s2 = sin(4*pi*month/12),
         month_s3 = sin(6*pi*month/12),
         month_s4 = sin(8*pi*month/12),
         month_c1 = cos(2*pi*month/12),
         month_c2 = cos(4*pi*month/12),
         month_c3 = cos(6*pi*month/12),
         month_c4 = cos(8*pi*month/12),
         yday_s1 = sin(2*pi*yday/365),
         yday_s2 = sin(4*pi*yday/365),
         yday_s3 = sin(6*pi*yday/365),
         yday_s4 = sin(8*pi*yday/365),
         yday_c1 = cos(2*pi*yday/365),
         yday_c2 = cos(4*pi*yday/365),
         yday_c3 = cos(6*pi*yday/365),
         yday_c4 = cos(8*pi*yday/365)) %>%
  drop_na(c(ndvi_temporal, veg_herby, canopy_01, sl_)) %>% 
  dplyr::select(!(burst_:case_))

```

Filtering by animals that have more than a year of data

```{r}

buffalo_year_ids <- c(2005, 2014, 2018, 2022, 2024, 2154, 2158, 2327, 2354, 2387)
buffalo_CLR_year <- buffalo_CLR %>% filter(id %in% buffalo_year_ids)
unique(buffalo_CLR_year$id)

```

```{r}

ggplot() +
  geom_line(data = buffalo_CLR_year %>% filter(id == 2005), aes(x = month, y = month_s1), size = 0.1) +
  geom_line(data = buffalo_CLR_year %>% filter(id == 2005), aes(x = month, y = month_s2), size = 0.1) +
  geom_line(data = buffalo_CLR_year %>% filter(id == 2005), aes(x = month, y = month_s3), size = 0.1) +
  geom_line(data = buffalo_CLR_year %>% filter(id == 2005), aes(x = month, y = month_s4), size = 0.1) +
  geom_line(data = buffalo_CLR_year %>% filter(id == 2005), aes(x = month, y = month_c1), size = 0.1) +
  geom_line(data = buffalo_CLR_year %>% filter(id == 2005), aes(x = month, y = month_c2), size = 0.1) +
  geom_line(data = buffalo_CLR_year %>% filter(id == 2005), aes(x = month, y = month_c3), size = 0.1) +
  geom_line(data = buffalo_CLR_year %>% filter(id == 2005), aes(x = month, y = month_c4), size = 0.1) +
  theme_classic()

ggplot() +
  geom_line(data = buffalo_CLR_year %>% filter(id == 2005), aes(x = yday, y = yday_s1), size = 0.1) +
  geom_line(data = buffalo_CLR_year %>% filter(id == 2005), aes(x = yday, y = yday_s2), size = 0.1) +
  geom_line(data = buffalo_CLR_year %>% filter(id == 2005), aes(x = yday, y = yday_s3), size = 0.1) +
  geom_line(data = buffalo_CLR_year %>% filter(id == 2005), aes(x = yday, y = yday_s4), size = 0.1) +
  geom_line(data = buffalo_CLR_year %>% filter(id == 2005), aes(x = yday, y = yday_c1), size = 0.1) +
  geom_line(data = buffalo_CLR_year %>% filter(id == 2005), aes(x = yday, y = yday_c2), size = 0.1) +
  geom_line(data = buffalo_CLR_year %>% filter(id == 2005), aes(x = yday, y = yday_c3), size = 0.1) +
  geom_line(data = buffalo_CLR_year %>% filter(id == 2005), aes(x = yday, y = yday_c4), size = 0.1) +
  theme_classic()

```



Tentative movement parameters from random sampling

```{r}

# gamma <- fit_distr(buffalo_all_steps_by_burst$sl_, "gamma")$params$shape
# vonmises <- fit_distr(buffalo_all_steps_by_burst$ta_, "vonmises")

# population tentative parameters
# gamma$params$shape # 0.438167
# gamma$params$scale # 534.3507
# vonmises$params$kappa # 0.1848126
# vonmises$params$mu # 0

movement_parameters <- read_csv("outputs/buffalo_parametric_indv_movement_params_20230208.csv")
movement_parameters_year <- movement_parameters %>% filter(id %in% buffalo_year_ids)

```


```{r}

coefs_df_function <- function(clogit_model) {
  
  return(data.frame("index" = 1:length(clogit_model$model$coefficients), 
                    "coef" = names(clogit_model$model$coefficients), 
                    "value" = clogit_model$model$coefficients, 
                    "se" = sqrt(diag(clogit_model$model$var)),
                    "weight" = 1/diag(clogit_model$model$var)))
}

```


```{r}

# hist(buffalo_subset$ndvi_temporal)
# hist(buffalo_subset$slope_scaled)
# hist(buffalo_subset$veg_herby)
# hist(buffalo_subset$canopy_scaled)

ndvi_above0_stack_scaled <- rast("mapping/cropped rasters/ndvi_above0_stack_scaled_20230210.tif")
# slope <- rast("mapping/cropped rasters/slope_raster.tif")
veg_herby <- rast("mapping/cropped rasters/veg_herby.tif")
# canopy_cover <- rast("mapping/cropped rasters/canopy_cover.tif")
slope_scaled <- rast("mapping/cropped rasters/slope_scaled_by_buffalo_data_20230210.tif")
canopy_scaled <- rast("mapping/cropped rasters/canopy_scaled_by_buffalo_data_20230210.tif")

water_dist_stack <- rast("mapping/cropped rasters/water_dist_stack_20230207.tif")
plot(water_dist_stack)
# water_dist_rasters <- raster::stack(water_dist_stack)
# water_dist_rasters <- raster::setZ(water_dist_rasters, terra::time(water_dist_stack))
# plot(water_dist_rasters)

years <- c(rep(2019, 7), rep(2018, 5))
months <- c(seq(1, 7, 1), seq(8, 12, 1))
day <- rep(1, 12)
dates <- make_datetime(year = years, month = months, day = day, tz = "Australia/Queensland")

```


## Conditional logistic regression model

```{r}

month_no <- 3

#########################

formula_clr <- y ~ 
  
  ndvi_scaled +
  I(ndvi_scaled^2) +
  
  # slope_scaled +
  # I(slope_scaled^2) +
  
  veg_herby +
  
  # elev_log +
  # I(elev_log^2) +
  
  # water_dist_log +
  # I(water_dist_log^2) +
  
  canopy_scaled +
  I(canopy_scaled^2) +
  
  sl +
  log_sl +
  cos_ta +
  strata(step_id)

#########################

# 2005 2014 2018 2022 2024 2154 2158 2327 2354 2387
buffalo_subset <- buffalo_CLR_year %>% filter(id == 2387 & month == month_no)

clogit_subset <- fit_clogit(buffalo_subset, formula = formula_clr)

# summary(clogit_subset)
# AIC(clogit_subset)
# clogit_subset$model$coefficients

#########################

clogit_coefs <-coefs_df_function(clogit_subset)

# clogit_model_coef_list <- map(clogit_model_list, coefs_df_function)

#########################

curve(clogit_coefs$value[clogit_coefs$coef == "ndvi_scaled"]*x + 
        clogit_coefs$value[clogit_coefs$coef == "I(ndvi_scaled^2)"]*x^2, from = -3, to = 3, n = 101)

# curve(clogit_coefs$value[clogit_coefs$coef == "slope_scaled"]*x + 
#         clogit_coefs$value[clogit_coefs$coef == "I(slope_scaled^2)"]*x^2, from = -3, to = 3, n = 101)

curve(clogit_coefs$value[clogit_coefs$coef == "veg_herby"]*x, from = -3, to = 3, n = 101)

curve(clogit_coefs$value[clogit_coefs$coef == "canopy_scaled"]*x + 
        clogit_coefs$value[clogit_coefs$coef == "I(canopy_scaled^2)"]*x^2, from = -3, to = 3, n = 101)

#########################

resources <- c(ndvi_above0_stack_scaled[[which(time(ndvi_above0_stack_scaled) == dates[month_no])]], 
               slope_scaled,
               veg_herby, 
               canopy_scaled)

plot(resources[[1]])
plot(resources[[2]])
plot(resources[[3]])
plot(resources[[4]])
plot(resources)

#########################

naive <- rast(resources)[[1]]
naive <- terra::setValues(naive, 0)
names(naive) <- "naive_predictions"

ndvi_naive <- resources[[1]] * clogit_coefs$value[clogit_coefs$coef == "ndvi_scaled"] + 
  resources[[1]]^2 * clogit_coefs$value[clogit_coefs$coef == "I(ndvi_scaled^2)"]
# slope_naive <- resources[[2]] * clogit_coefs$value[clogit_coefs$coef == "slope_scaled"] + 
#   resources[[2]]^2 * clogit_coefs$value[clogit_coefs$coef == "I(slope_scaled^2)"]
herby_naive <- resources[[3]] * clogit_coefs$value[clogit_coefs$coef == "veg_herby"]
canopy_naive <- resources[[4]] * clogit_coefs$value[clogit_coefs$coef == "canopy_scaled"] + 
  resources[[4]]^2 * clogit_coefs$value[clogit_coefs$coef == "I(canopy_scaled^2)"]

naive <- naive + ndvi_naive + herby_naive + canopy_naive # slope_naive + 
  
naive_norm <- exp(naive)/global(exp(naive), fun = "sum", na.rm = TRUE)[[1]]
terra::plot(naive_norm)
terra::plot(naive_norm)
points(buffalo_subset$x1, buffalo_subset$y1)

```


## Conditional logistic regression model - loop over individuals

```{r}

clogit_subset_list <- vector(mode = "list", length = length(buffalo_year_ids))

for(i in 1:length(buffalo_year_ids)) {

month_no <- i

#########################

formula_clr <- y ~ 
  
  ndvi_scaled +
  I(ndvi_scaled^2) +
  
  # slope_scaled +
  # I(slope_scaled^2) +
  
  veg_herby +
  
  # elev_log +
  # I(elev_log^2) +
  
  # water_dist_log +
  # I(water_dist_log^2) +
  
  canopy_scaled +
  I(canopy_scaled^2) +
  
  sl +
  log_sl +
  cos_ta +
  strata(step_id)

#########################

# 2005 2014 2018 2022 2024 2154 2158 2327 2354 2387
buffalo_subset <- buffalo_CLR_year %>% filter(id == buffalo_year_ids[[i]] & month == month_no)

clogit_subset_list[[i]] <- fit_clogit(buffalo_subset, formula = formula_clr)

# summary(clogit_subset)
# AIC(clogit_subset)
# clogit_subset$model$coefficients

}

#########################

clogit_model_coef_list <- map(clogit_subset_list, coefs_df_function)
clogit_coefs <- do.call(rbind, clogit_model_coef_list)
clogit_coefs$id <- rep(buffalo_year_ids, each = length(clogit_subset_list[[1]]$model$coefficients))
clogit_coefs

coef <- c()
mean <- c()
w_mean <- c()

for(i in 1:length(unique(clogit_coefs$coef))) {
  
  # i <- 1
  
  coef_i <- unique(clogit_coefs$coef)[i]

  coef <- c(coef, coef_i)
  
  mean <- c(mean, 
            mean(x = clogit_coefs[which(clogit_coefs$coef == coef_i),]$value))
  
  w_mean <- c(w_mean, 
            weighted.mean(x = clogit_coefs[which(clogit_coefs$coef == coef_i),]$value,
              w = clogit_coefs[which(clogit_coefs$coef == coef_i),]$weight))

}

clogit_weighted_means <- data.frame(coef, mean, w_mean)
clogit_weighted_means

#########################

curve(clogit_coefs$value[clogit_coefs$coef == "ndvi_scaled"]*x + 
        clogit_coefs$value[clogit_coefs$coef == "I(ndvi_scaled^2)"]*x^2, from = -3, to = 3, n = 101)

# curve(clogit_coefs$value[clogit_coefs$coef == "slope_scaled"]*x + 
#         clogit_coefs$value[clogit_coefs$coef == "I(slope_scaled^2)"]*x^2, from = -3, to = 3, n = 101)

curve(clogit_coefs$value[clogit_coefs$coef == "veg_herby"]*x, from = -3, to = 3, n = 101)

curve(clogit_coefs$value[clogit_coefs$coef == "canopy_scaled"]*x + 
        clogit_coefs$value[clogit_coefs$coef == "I(canopy_scaled^2)"]*x^2, from = -3, to = 3, n = 101)

#########################

resources <- c(ndvi_above0_stack_scaled[[which(time(ndvi_above0_stack_scaled) == dates[month_no])]], 
               slope_scaled,
               veg_herby, 
               canopy_scaled)

plot(resources[[1]])
plot(resources[[2]])
plot(resources[[3]])
plot(resources[[4]])
plot(resources)

#########################

naive <- rast(resources)[[1]]
naive <- terra::setValues(naive, 0)
names(naive) <- "naive_predictions"

ndvi_naive <- resources[[1]] * clogit_coefs$value[clogit_coefs$coef == "ndvi_scaled"] + 
  resources[[1]]^2 * clogit_coefs$value[clogit_coefs$coef == "I(ndvi_scaled^2)"]
# slope_naive <- resources[[2]] * clogit_coefs$value[clogit_coefs$coef == "slope_scaled"] + 
#   resources[[2]]^2 * clogit_coefs$value[clogit_coefs$coef == "I(slope_scaled^2)"]
herby_naive <- resources[[3]] * clogit_coefs$value[clogit_coefs$coef == "veg_herby"]
canopy_naive <- resources[[4]] * clogit_coefs$value[clogit_coefs$coef == "canopy_scaled"] + 
  resources[[4]]^2 * clogit_coefs$value[clogit_coefs$coef == "I(canopy_scaled^2)"]

naive <- naive + ndvi_naive + herby_naive + canopy_naive # slope_naive + 
  
naive_norm <- exp(naive)/global(exp(naive), fun = "sum", na.rm = TRUE)[[1]]
terra::plot(naive_norm)
terra::plot(naive_norm)
points(buffalo_subset$x1, buffalo_subset$y1)

```


# Using glmmTMB

```{r}

# from Palm et al - Palm, Eric C., Michael J. Suitor, Kyle Joly, Jim D. Herriges, Allicia P. Kelly, Dave Hervieux, Kelsey L. M. Russell, Torsten W. Bentzen, Nicholas C. Larter, and Mark Hebblewhite. 2022. “Increasing Fire Frequency and Severity Will Increase Habitat Loss for a Boreal Forest Indicator Species.” Ecological Applications: A Publication of the Ecological Society of America 32 (3): e2549.

m_bp_2_week_winter_no_FR <-
  glmmTMB(case_ ~ tree + I(tree^2) + TRI + I(TRI^2) + TPI + 
            lc_burn + lc_shrubs + lc_grass + lc_sparse + 
            lc_water + lc_other + lc_fen + log_step_length + 
            (1 | stratum) + (0 + tree | herd) + 
            (0 + TRI | herd) + (0 + lc_fen | herd) + (0 + lc_shrubs | herd) + 
            (0 + lc_other | herd) + (0 + lc_water | herd) + (0 + lc_grass | herd) + 
            (0 + lc_burn | herd/id), 
          data=bp_2_week_winter, family=poisson,
          map=list(theta=factor(c(NA, 1:9))),
          start=list(theta=c(log(1e3), rep(0, 9))))

```


Writing to our model - single month


```{r}

month_no <- 12
buffalo_month_subset <- buffalo_CLR_year %>% filter(month == month_no)

# using the full dataset
buffalo_month_subset <- buffalo_CLR_year

buffalo_month_subset <- buffalo_subset # from the SSF stan code

tic()
glmmTMB_model_month1 <- glmmTMB(y ~ -1 +
                           ndvi_scaled +
                            # I(ndvi_scaled^2) +
                             # poly(ndvi_scaled, degree = 2, raw = FALSE) +
                            # slope_scaled +
                            # I(slope_scaled^2) +
                             # veg_herby
                             herby_scaled +
                             # poly(veg_herby, degree = 1, raw = FALSE) +
                            # elev_log +
                            # I(elev_log^2) +
                            # water_dist_log +
                            # I(water_dist_log^2) +
                            canopy_scaled +
                            # I(canopy_scaled^2) +
                             # poly(canopy_scaled, degree = 2, raw = FALSE) +
                            sl +
                            log_sl +
                            cos_ta +
                             (1 | step_id) +
                           (0 + ndvi_scaled | id) +
                             # (0 + poly(ndvi_scaled, degree = 2, raw = FALSE) | id) +
                           # (0 + slope_scaled | id) +
                            (0 + herby_scaled | id) +
                             # (0 + poly(veg_herby, degree = 1, raw = FALSE) | id) +
                           (0 + canopy_scaled | id) +
                             # (0 + poly(canopy_scaled, degree = 2, raw = FALSE) | id),
                             (0 + sl | id) +
                           (0 + log_sl | id) +
                           (0 + cos_ta | id),
                           
                         data = buffalo_month_subset, 
                         family = poisson,
                         map = list(theta = factor(c(NA, 1:6))),
                         start = list(theta = c(log(1e3), rep(0,6)))
                         )
toc()

beep(sound = 2)

summary(glmmTMB_model_month1)


#########################


ndvi_above0_stack_scaled[[which(time(ndvi_above0_stack_scaled) == dates[month_no])]]

resources <- c(ndvi_above0_stack_scaled[[which(time(ndvi_above0_stack_scaled) == dates[month_no])]], 
               # slope_scaled,
               veg_herby, 
               canopy_scaled)


#########################


naive <- rast(resources)[[1]]
naive <- terra::setValues(naive, 0)
names(naive) <- "naive_predictions"

ndvi_lin <- glmmTMB_model_month1$fit$par[1]
ndvi_quad <- glmmTMB_model_month1$fit$par[2]
herby_lin <- glmmTMB_model_month1$fit$par[3]
canopy_lin <- glmmTMB_model_month1$fit$par[4]
canopy_quad <- glmmTMB_model_month1$fit$par[5]

x <- seq(-3, 3, 0.1)
y_ndvi <- x * ndvi_lin + x^2 * ndvi_quad
y_herby <- x * herby_lin
y_canopy <- x * canopy_lin + x^2 * canopy_quad

plot(x, y_ndvi)
plot(x, y_herby)
plot(x, y_canopy)

ndvi_naive <- resources[[1]] * ndvi_lin + resources[[1]]^2 * ndvi_quad
plot(ndvi_naive)

# slope_naive <- resources[[2]] * clogit_coefs$value[clogit_coefs$coef == "slope_scaled"] + 
#   resources[[2]]^2 * clogit_coefs$value[clogit_coefs$coef == "I(slope_scaled^2)"]

herby_naive <- resources[[2]] * herby_lin
plot(herby_naive)

canopy_naive <- resources[[3]] * canopy_lin + resources[[3]]^2 * canopy_quad
plot(canopy_naive)

naive <- naive + ndvi_naive + herby_naive + canopy_naive # slope_naive + 
naive_norm <- exp(naive)/global(exp(naive), fun = "sum", na.rm = TRUE)[[1]]

terra::plot(naive_norm)
terra::plot(naive_norm)
points(buffalo_month_subset$x1, buffalo_month_subset$y1)

```


## Looping over months

```{r}

glmmTMB_model_list <- vector(mode = "list", length = 12)
coefs_list <- vector(mode = "list", length = 12)
naive_norm_list <- vector(mode = "list", length = 12)

for(i in 1:12) {

month_no <- i

buffalo_month_subset <- buffalo_CLR_year %>% filter(month == month_no)

# tic()
glmmTMB_model_list[[i]] <- glmmTMB(y ~ -1 +
                           ndvi_scaled +
                            I(ndvi_scaled^2) +
                             # poly(ndvi_scaled, degree = 2, raw = FALSE) +
                            # slope_scaled +
                            # I(slope_scaled^2) +
                             veg_herby +
                             # poly(veg_herby, degree = 1, raw = FALSE) +
                            # elev_log +
                            # I(elev_log^2) +
                            # water_dist_log +
                            # I(water_dist_log^2) +
                            canopy_scaled +
                            I(canopy_scaled^2) +
                             # poly(canopy_scaled, degree = 2, raw = FALSE) +
                            sl +
                            log_sl +
                            cos_ta +
                             (1 | step_id) +
                           (0 + ndvi_scaled | id) +
                             # (0 + poly(ndvi_scaled, degree = 2, raw = FALSE) | id) +
                           # (0 + slope_scaled | id) +
                             (0 + veg_herby | id) +
                             # (0 + poly(veg_herby, degree = 1, raw = FALSE) | id) +
                           (0 + canopy_scaled | id),
                             # (0 + poly(canopy_scaled, degree = 2, raw = FALSE) | id),
                         data = buffalo_month_subset, 
                         family = poisson,
                         map = list(theta = factor(c(NA, 1:3))),
                         start = list(theta = c(log(1e3), rep(0,3)))
                         )
# toc()

summary(glmmTMB_model_list[[i]])


#########################


# ndvi_above0_stack_scaled[[which(time(ndvi_above0_stack_scaled) == dates[month_no])]]

resources <- c(ndvi_above0_stack_scaled[[which(time(ndvi_above0_stack_scaled) == dates[month_no])]], 
               # slope_scaled,
               veg_herby, 
               canopy_scaled)


#########################


naive <- rast(resources)[[1]]
naive <- terra::setValues(naive, 0)
names(naive) <- "naive_predictions"

# coefs_list <- glmmTMB_model_month1$fit$par[1:8]
coefs_list[[i]] <- glmmTMB_model_list[[i]]$fit$par[1:8]

ndvi_lin <- glmmTMB_model_list[[i]]$fit$par[1]
ndvi_quad <- glmmTMB_model_list[[i]]$fit$par[2]
herby_lin <- glmmTMB_model_list[[i]]$fit$par[3]
canopy_lin <- glmmTMB_model_list[[i]]$fit$par[4]
canopy_quad <- glmmTMB_model_list[[i]]$fit$par[5]

x <- seq(-3, 3, 0.1)
y_ndvi <- x * ndvi_lin + x^2 * ndvi_quad
y_herby <- x * herby_lin
y_canopy <- x * canopy_lin + x^2 * canopy_quad

plot(x, y_ndvi)
plot(x, y_herby)
plot(x, y_canopy)

ndvi_naive <- resources[[1]] * ndvi_lin + resources[[1]]^2 * ndvi_quad
# plot(ndvi_naive)

# slope_naive <- resources[[2]] * clogit_coefs$value[clogit_coefs$coef == "slope_scaled"] + 
#   resources[[2]]^2 * clogit_coefs$value[clogit_coefs$coef == "I(slope_scaled^2)"]

herby_naive <- resources[[2]] * herby_lin
# plot(herby_naive)

canopy_naive <- resources[[3]] * canopy_lin + resources[[3]]^2 * canopy_quad
# plot(canopy_naive)

naive <- naive + ndvi_naive + herby_naive + canopy_naive # slope_naive + 
naive_norm_list[[i]] <- exp(naive)/global(exp(naive), fun = "sum", na.rm = TRUE)[[1]]

terra::plot(naive_norm_list[[i]])
terra::plot(naive_norm_list[[i]])
points(buffalo_month_subset$x1, buffalo_month_subset$y1)

}

```

```{r}

for(i in 1:12) {
  terra::plot(naive_norm_list[[i]])
  points(buffalo_month_subset$x1, buffalo_month_subset$y1)
}

```



```{r}

do.call(rbind, coefs_list)

model_no <- 1:12
coefs <- c("ndvi_scaled", "I(ndvi_scaled^2)", "veg_herby", "canopy_scaled", "I(canopy_scaled^2)", "sl", "log_sl", "cos_ta")

data.frame("month" = rep(1:12, each = 8), "coef" = rep(coefs, 12), "beta" = coefs_list)

```


```{r}

cor(buffalo_CLR_year$veg_herby, buffalo_CLR_year$canopy_scaled)

```


```{r}

month_no <- 7

buffalo_month_subset <- buffalo_CLR_year %>% filter(month == month_no)

# sum(is.na(buffalo_CLR_year$water_dist_log))

tic()
glmmTMB_model_month1 <- glmmTMB(y ~ -1 +
                           # ndvi_scaled +
                           #  I(ndvi_scaled^2) +
                             poly(ndvi_scaled, degree = 2, raw = FALSE) +
                            # slope_scaled +
                            # I(slope_scaled^2) +
                             # herby_scaled +
                             # poly(veg_herby, degree = 1, raw = FALSE) +
                            # elev_log +
                            # I(elev_log^2) +
                             # poly(water_dist, degree = 2, raw = FALSE) +
                             # poly(water_dist_log, degree = 2, raw = FALSE) +
                             # water_dist +
                            # water_dist_log +
                            # I(water_dist_log^2) +
                            # canopy_scaled +
                            # I(canopy_scaled^2) +
                             poly(canopy_scaled, degree = 2, raw = FALSE) +
                             
                             sl +
                             log_sl +
                             cos_ta +
                             
                            # poly(sl, degree = 1, raw = FALSE) +
                            # poly(log_sl, degree = 1, raw = FALSE) +
                            #  poly(cos_ta, degree = 1, raw = FALSE) +
                             
                             (1 | step_id) +
                           # (0 + ndvi_scaled | id) +
                             (0 + poly(ndvi_scaled, degree = 2, raw = FALSE) | id) +
                           # (0 + slope_scaled | id) +
                             # (0 + poly(veg_herby, degree = 1, raw = FALSE) | id) +
                             # (0 + poly(water_dist, degree = 2, raw = FALSE) | id) +
                           # (0 + canopy_scaled | id)
                             (0 + poly(canopy_scaled, degree = 2, raw = FALSE) | id),
                           
                         data = buffalo_month_subset, 
                         family = poisson,
                         map = list(theta = factor(c(NA, 1:6))),
                         start = list(theta = c(log(1e3), rep(0,6))) # 
                         )
toc()

summary(glmmTMB_model_month1)

```



```{r}

glmmTMB_model_month1$fit$par

```


```{r}

x <- seq(-5,5,0.1)
x_poly <- poly(x, 2)

# plot(x, x)
# points(x, x^2)
# plot(x, x_poly[,1])
# points(x, x_poly[,2])

# x_poly

ndvi_poly <- poly(buffalo_month_subset$ndvi_scaled, degree = 2, raw = FALSE)
herby_poly <- poly(buffalo_month_subset$veg_herby, degree = 1, raw = FALSE)
canopy_poly <- poly(buffalo_month_subset$canopy_scaled, degree = 2, raw = FALSE)

ndvi_poly[1,]
x_ndvi <- buffalo_month_subset$ndvi_scaled[1]

alpha_coef <- attributes(ndvi_poly)$coefs$alpha
norm2_coef <- attributes(ndvi_poly)$coefs$norm2

f0 <- 1 / sqrt(norm2_coef[2])
# f0

# recovering a single ndvi value (converting from original scale to poly scale)
f1 <- (x_ndvi - alpha_coef[1]) / sqrt(norm2_coef[3])
# f1

f2 <- ((x_ndvi - alpha_coef[2]) * sqrt(norm2_coef[3]) * f1 - norm2_coef[3] / 
         sqrt(norm2_coef[2]) * f0) / sqrt(norm2_coef[4])
# f2

# f0_function <- function(x) (x - alpha_coef[1]) / sqrt(norm2_coef[3])

# without any fitted model coefficients
curve(expr = (x - alpha_coef[1]) / sqrt(norm2_coef[3]),
      from = -3, to = 3, n = 101)

curve(expr = ((x - alpha_coef[2]) * sqrt(norm2_coef[3]) * (x - alpha_coef[1]) / sqrt(norm2_coef[3]) - norm2_coef[3] / 
         sqrt(norm2_coef[2]) * f0) / sqrt(norm2_coef[4]),
      from = -3, to = 3, n = 101)

```

```{r}

ndvi_poly <- poly(buffalo_month_subset$ndvi_scaled, degree = 2, raw = FALSE)
alpha_coef <- attributes(ndvi_poly)$coefs$alpha
norm2_coef <- attributes(ndvi_poly)$coefs$norm2

ndvi_scaled <- ndvi_above0_stack_scaled[[12 + month_no]]
names(ndvi_scaled) <- "ndvi_scaled"
ndvi_df <- as.data.frame(ndvi_scaled, cells = TRUE, na.rm = FALSE)

# predict(data.frame(ndvi_scaled = ndvi_df[,3]), model = ndvi_poly)

f0 <- 1 / sqrt(norm2_coef[2])
# f0

f1 <- (ndvi_df[,2] - alpha_coef[1]) / sqrt(norm2_coef[3])
# f1

f2 <- ((ndvi_df[,2] - alpha_coef[2]) * sqrt(norm2_coef[3]) * f1 - norm2_coef[3] / 
         sqrt(norm2_coef[2]) * f0) / sqrt(norm2_coef[4])
# f2

ndvi_lin <- glmmTMB_model_month1$fit$par[1]
ndvi_quad <- glmmTMB_model_month1$fit$par[2]

ndvi_df$pred <- (f1 * ndvi_lin) + (f2 * ndvi_quad)

ndvi_pred <- ndvi_scaled
ndvi_pred <- terra::setValues(ndvi_pred, ndvi_df$pred)
names(ndvi_pred) <- "ndvi_predictions"

plot(ndvi_pred)


# plotting coefficient curve in original scale

x <- seq(-6, 3, 0.1)

x0 <- 1 / sqrt(norm2_coef[2])
x0

x1 <- (x - alpha_coef[1]) / sqrt(norm2_coef[3])
# f1

x2 <- ((x - alpha_coef[2]) * sqrt(norm2_coef[3]) * x1 - norm2_coef[3] / 
         sqrt(norm2_coef[2]) * x0) / sqrt(norm2_coef[4])

y_ndvi <- (x1 * ndvi_lin) + (x2 * ndvi_quad)

hist(ndvi_scaled)
hist(buffalo_month_subset$ndvi_scaled)
plot(x, y_ndvi)

```

Herbaceous vegetation

```{r}

herby_poly <- poly(buffalo_month_subset$veg_herby, degree = 1, raw = FALSE)
alpha_coef <- attributes(herby_poly)$coefs$alpha
norm2_coef <- attributes(herby_poly)$coefs$norm2

veg_herby_df <- as.data.frame(veg_herby, cells = TRUE, na.rm = FALSE)

# predict(data.frame(ndvi_scaled = ndvi_df[,3]), model = ndvi_poly)

f0 <- 1 / sqrt(norm2_coef[2])
# f0

f1 <- (veg_herby_df[,2] - alpha_coef[1]) / sqrt(norm2_coef[3])
# f1

# f2 <- ((ndvi_df[,2] - alpha_coef[2]) * sqrt(norm2_coef[3]) * f1 - norm2_coef[3] / 
#          sqrt(norm2_coef[2]) * f0) / sqrt(norm2_coef[4])
# f2

herby_lin <- glmmTMB_model_month1$fit$par[3]

veg_herby_df$pred <- f1 * ndvi_lin

herby_pred <- veg_herby
herby_pred <- terra::setValues(herby_pred, veg_herby_df$pred)
names(herby_pred) <- "herby_predictions"

plot(herby_pred)


# plotting coefficient curve in original scale

x <- seq(-2, 2, 0.1)

x0 <- 1 / sqrt(norm2_coef[2])
x0

x1 <- (x - alpha_coef[1]) / sqrt(norm2_coef[3])
# f1

y_herby <- (x1 * herby_lin)

hist(veg_herby)
hist(buffalo_month_subset$veg_herby)
plot(x, y_herby)

```

Distance to water

```{r}

water_poly <- poly(buffalo_month_subset$water_dist, degree = 1, raw = FALSE)
alpha_coef <- attributes(water_poly)$coefs$alpha
norm2_coef <- attributes(water_poly)$coefs$norm2

water_dist <- water_dist_stack[[12 + month_no]]
names(water_dist) <- "water_dist"

water_dist_df <- as.data.frame(water_dist, cells = TRUE, na.rm = FALSE)

# predict(data.frame(ndvi_scaled = ndvi_df[,3]), model = ndvi_poly)

f0 <- 1 / sqrt(norm2_coef[2])
# f0

f1 <- (veg_herby_df[,2] - alpha_coef[1]) / sqrt(norm2_coef[3])
# f1

# f2 <- ((ndvi_df[,2] - alpha_coef[2]) * sqrt(norm2_coef[3]) * f1 - norm2_coef[3] / 
#          sqrt(norm2_coef[2]) * f0) / sqrt(norm2_coef[4])
# f2

herby_lin <- glmmTMB_model_month1$fit$par[3]

veg_herby_df$pred <- f1 * ndvi_lin

herby_pred <- veg_herby
herby_pred <- terra::setValues(herby_pred, veg_herby_df$pred)
names(herby_pred) <- "herby_predictions"

plot(herby_pred)


# plotting coefficient curve in original scale

x <- seq(-2, 2, 0.1)

x0 <- 1 / sqrt(norm2_coef[2])
x0

x1 <- (x - alpha_coef[1]) / sqrt(norm2_coef[3])
# f1

y_herby <- (x1 * herby_lin)

hist(veg_herby)
hist(buffalo_month_subset$veg_herby)
plot(x, y_herby)

```


```{r}

ndvi_poly <- poly(buffalo_month_subset$ndvi_scaled, degree = 2, raw = FALSE)
alpha_coef <- attributes(ndvi_poly)$coefs$alpha
norm2_coef <- attributes(ndvi_poly)$coefs$norm2

ndvi_scaled <- ndvi_above0_stack_scaled[[12 + month_no]]
names(ndvi_scaled) <- "ndvi_scaled"
ndvi_df <- as.data.frame(ndvi_scaled, cells = TRUE, na.rm = FALSE)

# predict(data.frame(ndvi_scaled = ndvi_df[,3]), model = ndvi_poly)

f0 <- 1 / sqrt(norm2_coef[2])
# f0

f1 <- (ndvi_df[,2] - alpha_coef[1]) / sqrt(norm2_coef[3])
# f1

f2 <- ((ndvi_df[,2] - alpha_coef[2]) * sqrt(norm2_coef[3]) * f1 - norm2_coef[3] / 
         sqrt(norm2_coef[2]) * f0) / sqrt(norm2_coef[4])
# f2

ndvi_lin <- glmmTMB_model_month1$fit$par[1]
ndvi_quad <- glmmTMB_model_month1$fit$par[2]

ndvi_df$pred <- (f1 * ndvi_lin) + (f2 * ndvi_quad)

ndvi_pred <- ndvi_scaled
ndvi_pred <- terra::setValues(ndvi_pred, ndvi_df$pred)
names(ndvi_pred) <- "ndvi_predictions"

plot(ndvi_pred)


# plotting coefficient curve in original scale

x <- seq(-6, 3, 0.1)

x0 <- 1 / sqrt(norm2_coef[2])
x0

x1 <- (x - alpha_coef[1]) / sqrt(norm2_coef[3])
# f1

x2 <- ((x - alpha_coef[2]) * sqrt(norm2_coef[3]) * x1 - norm2_coef[3] / 
         sqrt(norm2_coef[2]) * x0) / sqrt(norm2_coef[4])

y_ndvi <- (x1 * ndvi_lin) + (x2 * ndvi_quad)

hist(ndvi_scaled)
hist(buffalo_month_subset$ndvi_scaled)
plot(x, y_ndvi)

```



```{r}

canopy_poly <- poly(buffalo_month_subset$canopy_scaled, degree = 2, raw = FALSE)
alpha_coef <- attributes(canopy_poly)$coefs$alpha
norm2_coef <- attributes(canopy_poly)$coefs$norm2

names(canopy_scaled) <- "canopy_scaled"
canopy_df <- as.data.frame(canopy_scaled, cells = TRUE, na.rm = FALSE)

# predict(data.frame(canopy_scaled = canopy_df[,3]), model = canopy_poly)

f0 <- 1 / sqrt(norm2_coef[2])
# f0

f1 <- (canopy_df[,2] - alpha_coef[1]) / sqrt(norm2_coef[3])
# f1

f2 <- ((canopy_df[,2] - alpha_coef[2]) * sqrt(norm2_coef[3]) * f1 - norm2_coef[3] / 
         sqrt(norm2_coef[2]) * f0) / sqrt(norm2_coef[4])
# f2

# canopy_lin <- glmmTMB_model_month1$fit$par[4]
# canopy_quad <- glmmTMB_model_month1$fit$par[5]

canopy_lin <- glmmTMB_model_month1$fit$par[3]
canopy_quad <- glmmTMB_model_month1$fit$par[4]

canopy_df$pred <- (f1 * canopy_lin) + (f2 * canopy_quad)

canopy_pred <- canopy_scaled
canopy_pred <- terra::setValues(canopy_pred, canopy_df$pred)
names(canopy_pred) <- "canopy_predictions"

plot(canopy_pred)


# plotting coefficient curve in original scale

x <- seq(-2.5, 2.5, 0.1)

x0 <- 1 / sqrt(norm2_coef[2])
x0

x1 <- (x - alpha_coef[1]) / sqrt(norm2_coef[3])
# f1

x2 <- ((x - alpha_coef[2]) * sqrt(norm2_coef[3]) * x1 - norm2_coef[3] / 
         sqrt(norm2_coef[2]) * x0) / sqrt(norm2_coef[4])

y_canopy <- (x1 * canopy_lin) + (x2 * canopy_quad)

hist(canopy_scaled)
hist(buffalo_month_subset$canopy_scaled)
plot(x, y_canopy)

```

Generating naive prediction map

```{r}

naive_pred <- exp(ndvi_pred + canopy_pred) # + herby_pred 
names(naive_pred) <- "naive_pred"

plot(naive_pred)
plot(naive_pred)
points(buffalo_month_subset$x1, buffalo_month_subset$y1)

```

crop raster to smaller extent

```{r}

min(buffalo_CLR_year$x1)
max(buffalo_CLR_year$x1)
min(buffalo_CLR_year$y1)
max(buffalo_CLR_year$y1)

xmin <- 5000
xmax <- 55000
ymin <- -1458000
ymax <- -1412000

template_raster <- rast(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, resolution = 25)

naive_pred_cropped <- terra::crop(naive_pred, template_raster)

plot(naive_pred_cropped)
plot(naive_pred_cropped)
points(buffalo_month_subset$x1, buffalo_month_subset$y1)

```


# Boyce index

```{r}

ecospat.boyce2 <- function (fit, obs, nclass = 0, window.w = "default", res = 100, 
  PEplot = TRUE, rm.duplicate = TRUE, method = "spearman") 
{
  boycei <- function(interval, obs, fit) {
    pi <- sum(as.numeric(obs >= interval[1] & obs <= interval[2]))/length(obs)
    ei <- sum(as.numeric(fit >= interval[1] & fit <= interval[2]))/length(fit)
    return(round(pi/ei, 10))
  }
  if (inherits(fit, "RasterLayer")) {
    if (is.data.frame(obs) || is.matrix(obs)) {
      obs <- raster::extract(fit, obs)
    }
    fit <- getValues(fit)
    fit <- fit[!is.na(fit)]
    obs <- obs[!is.na(obs)]
  }
  mini <- min(fit, obs)
  maxi <- max(fit, obs)
  if (length(nclass) == 1) {
    if (nclass == 0) {
      if (window.w == "default") {
        window.w <- (max(fit) - min(fit))/10
      }
      vec.mov <- seq(from = mini, to = maxi - window.w, 
        by = (maxi - mini - window.w)/res)
      vec.mov[res + 1] <- vec.mov[res + 1] + 1
      interval <- cbind(vec.mov, vec.mov + window.w)
    }
    else {
      vec.mov <- seq(from = mini, to = maxi, by = (maxi - 
        mini)/nclass)
      interval <- cbind(vec.mov, c(vec.mov[-1], maxi))
    }
  }
  else {
    vec.mov <- c(mini, sort(nclass[!nclass > maxi | nclass < 
      mini]))
    interval <- cbind(vec.mov, c(vec.mov[-1], maxi))
  }
  f <- apply(interval, 1, boycei, obs, fit)
  to.keep <- which(f != "NaN")
  f <- f[to.keep]
  if (length(f) < 2) {
    b <- NA
  }
  else {
    r <- 1:length(f)
    if (rm.duplicate == TRUE) {
      r <- c(1:length(f))[f != c(f[-1], TRUE)]
    }
    b <- cor(f[r], vec.mov[to.keep][r], method = method)
  }
  HS <- apply(interval, 1, sum)/2
  if (length(nclass) == 1 & nclass == 0) {
    HS[length(HS)] <- HS[length(HS)] - 1
  }
  HS <- HS[to.keep]
  if (PEplot == TRUE) {
    plot(HS, f, xlab = "Habitat suitability", ylab = "Predicted/Expected ratio", 
      col = "grey", cex = 0.75)
    points(HS[r], f[r], pch = 19, cex = 0.75)
  }
  return(list(F.ratio = f, cor = round(b, 3), HS = HS))
}

```


setting up the buffalo observation data for the relevant month

```{r}

# buffalo_obs <- buffalo_CLR_year %>% 
#   filter(y == 1 & month == 1 & x1 > xmin & x1 < xmax & y1 > ymin & y1 < ymax) %>% 
#   transmute(x = x1 - xmin, y = y1 - ymin)

buffalo_obs <- buffalo_CLR_year %>% filter(y == 1 & month == month_no) %>% 
  transmute(x = x1, y = y1)

```



```{r}

# naive_raster <- raster(naive_pred)
naive_raster <- raster(naive_pred_cropped)

# plot(naive_pred)
# plot(naive_raster)

# plot(naive_cropped_raster[[1]])

# boyce_naive <- ecospat.boyce2(naive_cropped_raster[[1]], buffalo_obs,
#                               # method = "spearman"
#                               method = "pearson",
#                               # method = "kendall"
#                               )

boyce_naive <- ecospat.boyce2(naive_raster, buffalo_obs,
                              method = "spearman"
                              # method = "pearson",
                              # method = "kendall"
                              )

boyce_naive

```



```{r}

naive_BI_scaled <- data.frame("method" = rep("naive", length(boyce_naive[[1]])), 
           "F.ratio" = boyce_naive[[1]],
           "HS" = (boyce_naive[[3]] - min(boyce_naive[[3]])) / 
             (max(boyce_naive[[3]]) - min(boyce_naive[[3]])))


plot(template_raster)
# template_raster <- raster::raster(template_raster)

boyce_sim <- ecospat.boyce2(template_raster, buffalo_obs,
                            method = "pearson")

boyce_sim

sim_BI_scaled <- data.frame("method" = rep("naive", length(boyce_sim[[1]])), 
           "F.ratio" = boyce_sim[[1]],
           "HS" = (boyce_sim[[3]] - min(boyce_sim[[3]])) / 
             (max(boyce_sim[[3]]) - min(boyce_sim[[3]])))

```


```{r}

ggplot() +
  geom_point(data = naive_BI_scaled, aes(x = HS, y = F.ratio), colour = "skyblue") +
  geom_path(data = naive_BI_scaled, aes(x = HS, y = F.ratio), colour = "skyblue") +
  geom_point(data = sim_BI_scaled, aes(x = HS, y = F.ratio), colour = "orange") +
  geom_path(data = sim_BI_scaled, aes(x = HS, y = F.ratio), colour = "orange") +
  theme_classic()

ggsave(paste0("outputs/plots/bcrw_simulations/boyce_index_naive_sims_small_extent_month", month_n, "_20230228.png"),
width=150, height=90, units="mm", dpi = 300)
# comment out geom_point size
ggsave(paste0("outputs/plots/bcrw_simulations/boyce_index_naive_sims_small_extent_month", month_n, "_pres_20230228.png"),
width=300, height=180, units="mm", dpi = 300, scale = 1)

```



### Quadratics with harmonics

