---
title: "SSF CLR fitting and prediction"
author: "Scott Forrest"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Import packages

```{r message = FALSE}

options(scipen=999)

# install.packages("clogitLasso")

library(tidyverse)
packages <- c("lubridate", "survival", "terra", "raster", "tictoc", "TwoStepCLogit", "ecospat", "beepr", "clogitL1", "clogitLasso")
walk(packages, require, character.only = T)

```

Import habitat layers 

```{r}

# xmin <- round(min(buffalo_CLR_year_harmonics$x2), digits = -2)
# xmax <- round(max(buffalo_CLR_year_harmonics$x2), digits = -2)
# ymin <- round(min(buffalo_CLR_year_harmonics$y2), digits = -2)
# ymax <- round(max(buffalo_CLR_year_harmonics$y2), digits = -2)
# 
# # scaled rasters
# ndvi_stack_scaled_cropped <- rast("mapping/cropped rasters/ndvi_projected_watermask_scaled_cropped_20230612.tif")
# canopy_scaled_cropped <- rast("mapping/cropped rasters/canopy_scaled_cropped_by_buffalo_data_20230210.tif")
# herby_scaled_cropped <- rast("mapping/cropped rasters/herby_scaled_by_buffalo_data_cropped_20230210.tif")
# elev_scaled_cropped <- rast("mapping/cropped rasters/elev_scaled_by_buffalo_data_cropped_20230613.tif")
# 
# # to set the origin at (0,0)
# ext(ndvi_stack_scaled_cropped) <- c(xmin - xmin, xmax - xmin, ymin - ymin, ymax - ymin)
# ext(canopy_scaled_cropped) <- c(xmin - xmin, xmax - xmin, ymin - ymin, ymax - ymin)
# ext(herby_scaled_cropped) <- c(xmin - xmin, xmax - xmin, ymin - ymin, ymax - ymin)
# ext(elev_scaled_cropped) <- c(xmin - xmin, xmax - xmin, ymin - ymin, ymax - ymin)
                              
```

For unscaled rasters

```{r}

# unscaled rasters
ndvi_stack <- rast("mapping/cropped rasters/ndvi_GEE_projected_watermask20230207.tif")
canopy <- rast("mapping/cropped rasters/canopy_cover.tif")
herby <- rast("mapping/cropped rasters/veg_herby.tif")
elev <- rast("mapping/cropped rasters/DEM_H_raster.tif")

crop_extent <- ext(xmin, xmax, ymin, ymax)

ndvi_stack_cropped <- terra::crop(ndvi_stack, crop_extent)
canopy_cropped <- terra::crop(canopy, crop_extent)
herby_cropped <- terra::crop(herby, crop_extent)
elev_cropped <- terra::crop(elev, crop_extent)

# to set the origin at (0,0)
ext(ndvi_stack_cropped) <- c(xmin - xmin, xmax - xmin, ymin - ymin, ymax - ymin)
ext(canopy_cropped) <- c(xmin - xmin, xmax - xmin, ymin - ymin, ymax - ymin)
ext(herby_cropped) <- c(xmin - xmin, xmax - xmin, ymin - ymin, ymax - ymin)
ext(elev_cropped) <- c(xmin - xmin, xmax - xmin, ymin - ymin, ymax - ymin)

# plot(ndvi_stack_cropped)
# plot(canopy_cropped)
# plot(herby_cropped)
# plot(elev_cropped)

```

Using the lasso approach

```{r}

# set.seed(145)
# data parameters
K = 10 # number of strata
n = 5 # number in strata
m = 2 # cases per stratum
p = 20 # predictors

# generate data
y = rep(c(rep(1, m), rep(0, n-m)), K)
X = matrix (rnorm(K*n*p, 0, 1), ncol = p) # pure noise
strata = sort(rep(1:K, n))

par(mfrow = c(1,2))
# fit the conditional logistic model
clObj = clogitL1(y=y, x=X, strata)
plot(clObj, logX=TRUE)

# cross validation
clcvObj = cv.clogitL1(clObj)
plot(clcvObj)

```

Removing negative values from elevation

```{r}

buffalo_CLR_year_harmonics <- buffalo_CLR_year_harmonics %>% mutate(elev_0 = ifelse(DEM_H_end < 1, 1, DEM_H_end),
                                                                    elev_log = log(elev_0))
hist(buffalo_CLR_year_harmonics$DEM_H_end)
hist(buffalo_CLR_year_harmonics$elev_0)
hist(buffalo_CLR_year_harmonics$elev_log)

sum(is.na(buffalo_CLR_year_harmonics$elev_log))

```



```{r}

buffalo_ids <- unique(buffalo_CLR_year_harmonics$id)
which_buffalo <- 3

buffalo_data <- buffalo_CLR_year_harmonics %>% filter(id == buffalo_ids[which_buffalo]) %>% drop_na(cos_ta)

buffalo_lasso_unscaled <- buffalo_data %>% mutate(
  
  ndvi = ndvi_temporal,
  ndvi_s1 = ndvi_temporal * yday_s1,
  ndvi_s2 = ndvi_temporal * yday_s2,
  # ndvi_s3 = ndvi_temporal * yday_s3,
  # ndvi_s4 = ndvi_temporal * yday_s4,
  ndvi_c1 = ndvi_temporal * yday_c1,
  ndvi_c2 = ndvi_temporal * yday_c2,
  # ndvi_c3 = ndvi_temporal * yday_c3,
  # ndvi_c4 = ndvi_temporal * yday_c4,
  
  ndvi_sq = ndvi_temporal ^ 2,
  ndvi_sq_s1 = (ndvi_temporal ^ 2) * yday_s1,
  ndvi_sq_s2 = (ndvi_temporal ^ 2) * yday_s2,
  # ndvi_sq_s3 = (ndvi_temporal ^ 2) * yday_s3,
  # ndvi_sq_s4 = (ndvi_temporal ^ 2) * yday_s4,
  ndvi_sq_c1 = (ndvi_temporal ^ 2) * yday_c1,
  ndvi_sq_c2 = (ndvi_temporal ^ 2) * yday_c2,
  # ndvi_sq_c3 = (ndvi_temporal ^ 2) * yday_c3,
  # ndvi_sq_c4 = (ndvi_temporal ^ 2) * yday_c4,
  
  canopy = canopy_cover,
  canopy_s1 = canopy_cover * yday_s1,
  canopy_s2 = canopy_cover * yday_s2,
  # canopy_s3 = canopy_cover * yday_s3,
  # canopy_s4 = canopy_cover * yday_s4,
  canopy_c1 = canopy_cover * yday_c1,
  canopy_c2 = canopy_cover * yday_c2,
  # canopy_c3 = canopy_cover * yday_c3,
  # canopy_c4 = canopy_cover * yday_c4,
  
  canopy_sq = canopy_cover ^ 2,
  canopy_sq_s1 = (canopy_cover ^ 2) * yday_s1,
  canopy_sq_s2 = (canopy_cover ^ 2) * yday_s2,
  # canopy_sq_s3 = (canopy_cover ^ 2) * yday_s3,
  # canopy_sq_s4 = (canopy_cover ^ 2) * yday_s4,
  canopy_sq_c1 = (canopy_cover ^ 2) * yday_c1,
  canopy_sq_c2 = (canopy_cover ^ 2) * yday_c2,
  # canopy_sq_c3 = (canopy_cover ^ 2) * yday_c3,
  # canopy_sq_c4 = (canopy_cover ^ 2) * yday_c4,
  
  herby = veg_herby,
  herby_s1 = veg_herby * yday_s1,
  herby_s2 = veg_herby * yday_s2,
  # herby_s3 = veg_herby * yday_s3,
  # herby_s4 = veg_herby * yday_s4,
  herby_c1 = veg_herby * yday_c1,
  herby_c2 = veg_herby * yday_c2,
  # herby_c3 = veg_herby * yday_c3,
  # herby_c4 = veg_herby * yday_c4,
  
  elev = elev_log,
  elev_s1 = elev_log * yday_s1,
  elev_s2 = elev_log * yday_s2,
  # elev_s3 = elev_log * yday_s3,
  elev_c1 = elev_log * yday_c1,
  elev_c2 = elev_log * yday_c2,
  # elev_c3 = elev_log * yday_c3,
  # 
  elev_log_sq = (elev_log ^ 2),
  elev_log_sq_s1 = (elev_log ^ 2) * yday_s1,
  elev_log_sq_s2 = (elev_log ^ 2) * yday_s2,
  # elev_log_sq_s3 = (elev_log ^ 2) * yday_s3,
  elev_log_sq_c1 = (elev_log ^ 2) * yday_c1,
  elev_log_sq_c2 = (elev_log ^ 2) * yday_c2,
  # elev_log_sq_c3 = (elev_log ^ 2) * yday_c3,
  
  step_l = sl,
  step_l_s1 = sl * yday_s1,
  step_l_s2 = sl * yday_s2,
  # step_l_s3 = sl * yday_s3,
  # step_l_s4 = sl * yday_s4,
  step_l_c1 = sl * yday_c1,
  step_l_c2 = sl * yday_c2,
  # step_l_c3 = sl * yday_c3,
  # step_l_c4 = sl * yday_c4,

  log_step_l = log_sl,
  log_step_l_s1 = log_sl * yday_s1,
  log_step_l_s2 = log_sl * yday_s2,
  # log_step_l_s3 = log_sl * yday_s3,
  # log_step_l_s4 = log_sl * yday_s4,
  log_step_l_c1 = log_sl * yday_c1,
  log_step_l_c2 = log_sl * yday_c2,
  # log_step_l_c3 = log_sl * yday_c3,
  # log_step_l_c4 = log_sl * yday_c4,

  cos_turn_a = cos_ta,
  cos_turn_a_s1 = cos_ta * yday_s1,
  cos_turn_a_s2 = cos_ta * yday_s2,
  # cos_turn_a_s3 = cos_ta * yday_s3,
  # cos_turn_a_s4 = cos_ta * yday_s4,
  cos_turn_a_c1 = cos_ta * yday_c1,
  cos_turn_a_c2 = cos_ta * yday_c2,
  # cos_turn_a_c3 = cos_ta * yday_c3,
  # cos_turn_a_c4 = cos_ta * yday_c4,
  
  .keep = "none"
  
)

# mean_vals <- apply(buffalo_lasso_unscaled, 2, mean)
# sd_vals <- apply(buffalo_lasso_unscaled, 2, sd)
# scaling_attributes <- data.frame(variable = names(buffalo_lasso_unscaled), mean = mean_vals, sd = sd_vals)

buffalo_lasso_scaled <- scale(buffalo_lasso_unscaled)

mean_vals <- attr(buffalo_lasso_scaled, "scaled:center")
sd_vals <- attr(buffalo_lasso_scaled, "scaled:scale")
scaling_attributes <- data.frame(variable = names(buffalo_lasso_unscaled), mean = mean_vals, sd = sd_vals)

# generate data
y = buffalo_data$y
X = data.matrix(buffalo_lasso_scaled)
strata = buffalo_data$step_id

# attr(X, "dimnames") <- NULL

# par(mfrow = c(1,2))
# fit the conditional logistic model
tic()
cl_buffalo = clogitL1(y=y, 
                      x=X, 
                      strata, 
                       # numLambda = 100,
                       # minLambdaRatio = 1e-8
                      alpha = 1)
toc()
plot(cl_buffalo, logX=TRUE)

# cross validation
tic()
clcv_buffalo = cv.clogitL1(cl_buffalo, numFolds = 4)
toc()
plot(clcv_buffalo)

beep(sound = 2)

```

Just plotting 

```{r}

plot(cl_buffalo, logX=TRUE)
plot(clcv_buffalo)

# summary(cl_buffalo)
summary(clcv_buffalo)

```


Fitting the same data with the non-penalised approach as a comparison - IT IS EXACTLY THE SAME

```{r}

clogit_df <- data.frame(y, strata, X)

clogit_buffalo <- clogit(formula = y ~ 
  
  ndvi +
  ndvi_s1 +
  ndvi_c1 +
  ndvi_s2 +
  ndvi_c2 +
  ndvi_s3 +
  ndvi_c3 +
  
  ndvi_sq +
  ndvi_sq_s1 +
  ndvi_sq_c1 +
  ndvi_sq_s2 +
  ndvi_sq_c2 +
  ndvi_sq_s3 +
  ndvi_sq_c3 +
  
  canopy +
  canopy_s1 +
  canopy_c1 +
  canopy_s2 +
  canopy_c2 +
  canopy_s3 +
  canopy_c3 +
  
  canopy_sq +
  canopy_sq_s1 +
  canopy_sq_c1 +
  canopy_sq_s2 +
  canopy_sq_c2 +
  canopy_sq_s3 +
  canopy_sq_c3 +
  
  herby +
  herby_s1 +
  herby_c1 +
  herby_s2 +
  herby_c2 +
  herby_s3 +
  herby_c3 +
  
  # elev_scaled +
  # yday_s1:elev_scaled +
  # yday_c1:elev_scaled +
  # yday_s2:elev_scaled +
  # yday_c2:elev_scaled +
  # yday_s3 +
  # yday_c3 +
  # 
  # elev_2 +
  # yday_s1:elev_2 +
  # yday_c1:elev_2 +
  # yday_s2:elev_2 +
  # yday_c2:elev_2 +
  # yday_s3 +
  # yday_c3 +
  
  step_l +
  step_l_s1 +
  step_l_c1 +
  step_l_s2 +
  step_l_c2 +
  step_l_s3 +
  step_l_c3 +
  
  log_step_l +
  log_step_l_s1 +
  log_step_l_c1 +
  log_step_l_s2 +
  log_step_l_c2 +
  log_step_l_s3 +
  log_step_l_c3 +
  
  cos_turn_a +
  cos_turn_a_s1 +
  cos_turn_a_c1 +
  cos_turn_a_s2 +
  cos_turn_a_c2 +
  cos_turn_a_s3 +
  cos_turn_a_c3 +
    
    strata(strata),
  
  data = clogit_df)

```

Summary

```{r}

clogit_buffalo

```

With clogitLasso package

```{r}

# y <- rep(c(1,rep(0,9)), 100)
# X <- matrix (rnorm(100000, 0, 1), ncol = 100) # pure noise
# strata <- sort(rep(1:100, 10))
# 
# # fit the conditional logistic model
# tic()
# classo_buffalo = clogitLasso(X=X, y=y, strata=strata)
# toc()
# plot(classo_buffalo)
# # plot(cl_buffalo2, logX=TRUE)
# 
# table(strata)[1]

```

Reconstructing LASSO coefficients

```{r}

yday <- seq(0,365,1)
yday_harmonics_df <- data.frame("linear_term" = rep(1, length(yday)),
                                "yday_s1" = sin(2*pi*yday/365),
                                "yday_s2" = sin(4*pi*yday/365),
                                # "yday_s3" = sin(6*pi*yday/365),
                                # "yday_s4" = sin(8*pi*yday/365),
                                "yday_c1" = cos(2*pi*yday/365),
                                "yday_c2" = cos(4*pi*yday/365))
                                # "yday_c3" = cos(6*pi*yday/365),
                                # "yday_c4" = cos(8*pi*yday/365))

# coefs <- attr(X, "dimnames")[[2]]
# coef_betas_CV <- summary(clcv_buffalo)$beta_minCV
# coef_betas_CV1se <- summary(clcv_buffalo)$beta_minCV1se
# coef_betas_unpen <- cl_buffalo$beta[101,]

# rescaling parameters back to natural scale
coefs_lasso <- data.frame("coefs" = attr(X, "dimnames")[[2]],
                          scaling_attributes,
                          "beta_unpen_scl" = cl_buffalo$beta[101,], 
                          "beta_CV_scl" = summary(clcv_buffalo)$beta_minCV,
                          "beta_CV1se_scl" = summary(clcv_buffalo)$beta_minCV1se)

coefs_lasso <- coefs_lasso %>% mutate(beta_unpen_nat = (beta_unpen_scl / sd),
                                      beta_CV_nat = (beta_CV_scl / sd),
                                      beta_CV1se_nat = (beta_CV1se_scl / sd)) #  + mean

# scaled
# model_fit <- "coef_betas_unpen_scl"
model_fit <- "beta_CV_scl"
# model_fit <- "beta_CV1se_scl"
# natural scale
# model_fit <- "coef_betas_unpen_nat"
# model_fit <- "beta_CV_nat"
# model_fit <- "beta_CV1se_nat"

harmonics_df <- data.frame("yday" = yday,
                           "ndvi" = as.numeric(coefs_lasso %>% dplyr::filter(grepl("ndvi", coefs) & !grepl("sq", coefs)) %>% 
                                                    dplyr::select(all_of(model_fit)) %>% unlist() %>% as.numeric() 
                                                  %*% t(as.matrix(yday_harmonics_df))),
                           "ndvi_2" = as.numeric(coefs_lasso %>% dplyr::filter(grepl("ndvi_sq", coefs)) %>% 
                                                    dplyr::select(model_fit) %>% unlist() %>% as.numeric() 
                                                  %*% t(as.matrix(yday_harmonics_df))),
                           "canopy" = as.numeric(coefs_lasso %>% dplyr::filter(grepl("canopy", coefs) & !grepl("sq", coefs)) %>% 
                                                    dplyr::select(model_fit) %>% unlist() %>% as.numeric() 
                                                  %*% t(as.matrix(yday_harmonics_df))),
                           "canopy_2" = as.numeric(coefs_lasso %>% dplyr::filter(grepl("canopy_sq", coefs)) %>% 
                                                    dplyr::select(model_fit) %>% unlist() %>% as.numeric() 
                                                  %*% t(as.matrix(yday_harmonics_df))),
                           "herby" = as.numeric(coefs_lasso %>% dplyr::filter(grepl("herby", coefs)) %>% 
                                                    dplyr::select(model_fit) %>% unlist() %>% as.numeric() 
                                                  %*% t(as.matrix(yday_harmonics_df))),
                           "sl" = as.numeric(coefs_lasso %>% dplyr::filter(grepl("step", coefs) & !grepl("log", coefs)) %>% 
                                                    dplyr::select(model_fit) %>% unlist() %>% as.numeric() 
                                                  %*% t(as.matrix(yday_harmonics_df))),
                           "log_sl" = as.numeric(coefs_lasso %>% dplyr::filter(grepl("log_step", coefs)) %>% 
                                                    dplyr::select(model_fit) %>% unlist() %>% as.numeric() 
                                                  %*% t(as.matrix(yday_harmonics_df))),
                           "cos_ta" = as.numeric(coefs_lasso %>% dplyr::filter(grepl("cos", coefs)) %>% 
                                                    dplyr::select(model_fit) %>% unlist() %>% as.numeric() 
                                                  %*% t(as.matrix(yday_harmonics_df))))

```

Movement parameters

```{r}

# for individuals
# movement_parameters <- read_csv("outputs/buffalo_parametric_indv_movement_params_20230208.csv")
# movement_parameters_year <- movement_parameters %>% filter(id %in% buffalo_year_ids)

# for the population
# gamma$params$shape # 0.438167
# gamma$params$scale # 534.3507
# vonmises$params$kappa # 0.1848126

tentative_shape <- 0.438167
tentative_scale <- 534.3507
tentative_kappa <- 0.1848126

yday_coefs <- harmonics_df %>% mutate(shape = tentative_shape + log_sl,
                             scale = 1/((1/tentative_scale) - sl),
                             kappa = tentative_kappa + cos_ta)

# write_csv(harmonic_coefs_df, "outputs/harmonic_coefs_df_with_movement_params_ndvi_canopy_quads_20230519.csv")

# turning into a long data frame
yday_coefs_long <- pivot_longer(yday_coefs, cols = !1, names_to = "coef")

# write_csv(harmonic_coefs, "outputs/harmonic_coefs_df_long_with_movement_params_ndvi_canopy_quads_20230519.csv")

```



```{r}

# coefs <- unique(yday_coefs_long$coef)
# coef_titles <- c("NDVI", "Slope", "Herbaceous vegetation", "Canopy cover", "Step length", "log(Step length)", "cos(Turning angle)")

ggplot() +
    geom_path(data = yday_coefs_long %>%
                filter(coef %in% unique(yday_coefs_long$coef)[1:5]),
              aes(x = yday, y = value, colour = coef)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    scale_y_continuous(expression(Time-varying~parameter~values~beta)) +
    scale_x_continuous("Day of the year") +
    scale_color_discrete("Estimate") +
    # ggtitle(coef_titles[i]) +
    theme_classic() +
    theme(legend.position = "bottom")

# ggsave(paste("outputs/plots/clr_fitting/clr_3harmonics_all_quads_20230519.png", sep = ""),
#   width=150, height=90, units="mm", dpi = 300)
# ggsave(paste("outputs/plots/clr_fitting/clr_harmonic_all_pres_20230208.png", sep = ""),
#   width=300, height=180, units="mm", dpi = 300)

ggplot() +
    geom_path(data = yday_coefs_long %>% 
                filter(coef %in% unique(yday_coefs_long$coef)[c(9,11)]), 
              aes(x = yday, y = value, colour = coef)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    scale_y_continuous(expression(beta)) +
    scale_x_continuous("Day of the year") +
    scale_color_discrete("Estimate") +
    # ggtitle(coef_titles[i]) +
    theme_classic() +
    theme(legend.position = "bottom")

ggplot() +
    geom_path(data = yday_coefs_long %>% 
                filter(coef %in% unique(yday_coefs_long$coef)[10]), 
              aes(x = yday, y = value, colour = coef)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    scale_y_continuous(expression(beta)) +
    scale_x_continuous("Day of the year") +
    scale_color_discrete("Estimate") +
    # ggtitle(coef_titles[i]) +
    theme_classic() +
    theme(legend.position = "bottom")

```


```{r temporal naive}

years <- c(rep(2019, 7), rep(2018, 5))
months <- c(seq(1, 7, 1), seq(8, 12, 1))
day <- rep(1, 12)
dates <- make_datetime(year = years, month = months, day = day, tz = "Australia/Queensland")

mid_month <- c(16, 47, 75, 106, 136, 167, 197, 228, 259, 289, 320, 350)

# get the coefficient at the middle of each month (15th day)
monthly_coefs <- data.frame("date" = dates,
                         "ndvi" = yday_coefs$ndvi[mid_month],
                         "ndvi_quad" = yday_coefs$ndvi_2[mid_month],
                         "canopy" = yday_coefs$canopy[mid_month],
                         "canopy_quad" = yday_coefs$canopy_2[mid_month],
                         "herby" = yday_coefs$herby[mid_month],
                         # "elev" = yday_coefs$elev[mid_month],
                         # "elev_quad" = yday_coefs$elev_quad[mid_month],
                         "shape" = yday_coefs$shape[mid_month],
                         "scale" = yday_coefs$scale[mid_month],
                         "kappa" = yday_coefs$kappa[mid_month])

# write_csv(monthly_coefs, "outputs/monthly_coefs_wide_yday_with_movement_params_20230530.csv") 

monthly_coefs_long <- monthly_coefs %>% pivot_longer(cols = !date)

# sanity check to ensure that the coefficients line up to the correct months
monthly_coefs_long %>% filter(!name %in% c("shape", "scale", "kappa")) %>% 
  ggplot(aes(x = month(date), y = value, colour = factor(name))) +
  geom_line() +
  scale_colour_viridis_d("Covariate") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_x_continuous(breaks = 1:12) +
  theme_classic()

```

# Predictions and cross-validation

## Naive prediction approach

```{r}

boyce_naive_list <- vector(mode = "list", length = 12)
naive_pred_stack <- c()
naive_norm_stack <- c()

for(i in 1:12) {

  # with scaled parameters
# resources <- c(ndvi_stack_scaled_cropped[[which(time(ndvi_stack_scaled_cropped) == monthly_coefs$date[i])]], 
#                canopy_scaled_cropped,
#                herby_scaled_cropped
#                # elev_scaled_cropped
#                )

# with natural scaled parameters
resources <- c(ndvi_stack_cropped[[which(time(ndvi_stack_cropped) == monthly_coefs$date[i])]], 
               canopy_cropped,
               herby_cropped
               # elev_scaled_cropped
               )

# ndvi
ndvi_lin <- resources[[1]]
ndvi_lin <- ndvi_lin * monthly_coefs$ndvi[[i]]
# plot(ndvi_lin)
ndvi_quad <- resources[[1]]
ndvi_quad <- (ndvi_quad ^ 2) * monthly_coefs$ndvi_quad[[i]] 
# plot(ndvi_quad)
ndvi_pred <- ndvi_lin + ndvi_quad
# plot(ndvi_pred)

# canopy cover 
canopy_lin <- resources[[2]]
canopy_lin <- canopy_lin * monthly_coefs$canopy[[i]]
# plot(canopy_lin)
canopy_quad <- resources[[2]]
canopy_quad <- (canopy_quad ^ 2) * monthly_coefs$canopy_quad[[i]] 
# plot(canopy_quad)
canopy_pred <- canopy_lin + canopy_quad
# plot(canopy_pred)

# herby
herby_lin <- resources[[3]]
herby_pred <- herby_lin * monthly_coefs$herby[[i]] 
# plot(herby_pred)

# # elev
# elev_lin <- resources[[4]]
# elev_lin <- elev_lin * monthly_coefs$elev[[i]] 
# # plot(elev_pred)
# elev_quad <- resources[[4]]
# elev_quad <- (elev_quad ^ 2) * monthly_coefs$elev_quad[[i]] 
# # plot(elev_quad)
# elev_pred <- elev_lin + elev_quad
# # plot(elev_pred)

# combining
naive_pred <- exp(ndvi_pred + 
                    canopy_pred + 
                    herby_pred #+ 
                    # elev_pred
                  )

naive_norm <- naive_pred / global(naive_pred, fun = "sum", na.rm = TRUE)[[1]]

plot(naive_norm)
points(as.numeric(buffalo_CLR_year %>% filter(y == 1 & month == i) %>% dplyr::select(x1) %>% unlist()), 
       as.numeric(buffalo_CLR_year %>% filter(y == 1 & month == i) %>% dplyr::select(y1) %>% unlist()))

naive_pred_stack <- c(naive_pred_stack, naive_pred)
naive_norm_stack <- c(naive_norm_stack, naive_norm)

# setting up the buffalo observation data for the relevant month
buffalo_obs <- buffalo_CLR_year %>% filter(y == 1 & month == i) %>% 
  transmute(x = x1 - xmin, y = y1 - ymin)

naive_raster <- raster(naive_norm)

boyce_naive_list[[i]] <- ecospat.boyce2(naive_raster, buffalo_obs,
                              method = "spearman"
                              # method = "pearson",
                              # method = "kendall"
                              )

boyce_naive_list[[i]]

}

naive_pred_lasso <- naive_pred_stack
naive_norm_lasso <- naive_norm_stack
boyce_naive_list_lasso <- boyce_naive_list

beep(sound = 2)

```


```{r}

for(i in 1:12) {
  plot(naive_pred_stack[[i]])
}

for(i in 1:12) {
  # plot(naive_pred_stack[[i]])
  plot(naive_pred_stack[[i]])
  points(as.numeric(buffalo_CLR_year %>% filter(y == 1 & month == i) %>% dplyr::select(x1) %>% unlist()) - xmin, 
       as.numeric(buffalo_CLR_year %>% filter(y == 1 & month == i) %>% dplyr::select(y1) %>% unlist()) - ymin)
}

```


```{r}

spearman_cor <- c()

for(i in 1:12) {
  print(boyce_naive_list[[i]]$cor)
  spearman_cor[i] <- boyce_naive_list[[i]]$cor
}

mean(spearman_cor)
sd(spearman_cor)

```

# Using the Barnett-Moorcroft approximation prediction approach with Monte Carlo approximation

Here we use the Barnett-Moorcroft approximation to estimate the UD, sampling from the movement kernel to approximate the integral in the numerator of:

BM equation

```{r}

naive_preds <- rast(naive_pred_stack)

Beta_Z_z_list  <- vector(mode = "list", length = 12)
MCBM_preds_list <- vector(mode = "list", length = 12)

n_proposals <- 10
grid_res <- terra::res(naive_preds)[1]
rows <- terra::nrow(naive_preds)
cols <- terra::ncol(naive_preds)
xmax <- ext(naive_preds[[1]])[2] 
ymax <- ext(naive_preds[[1]])[4]

for(month_no in 1:12) {
  
  # for testing the loop
  # month_no <- 9

tic()

# Create grid of x and y points
x_points <- rep(1:cols, each = rows)
y_points <- rep(1:rows, times = cols)

# Generate random angles and lengths
ta <- runif(n_proposals * length(x_points), min = -pi, max = pi)
sl <- rgamma(n_proposals * length(x_points), shape = monthly_coefs$shape[month_no], scale = monthly_coefs$scale[month_no])

# Calculate proposal points
x_proposal <- ((-(grid_res/2) + x_points * grid_res) + sl * sin(ta)) %% xmax
y_proposal <- ((-(grid_res/2) + y_points * grid_res) + sl * cos(ta)) %% ymax

# exp(beta * Z(x))
# plot(naive_ud_cropped[[i]])

# exp(beta * Z(z)) * psi(.) dz
Beta_Z_z_proposed <- terra::extract(naive_preds[[month_no]], cbind(x_proposal, y_proposal))[,1]
Beta_Z_z_array <- array(Beta_Z_z_proposed, dim = c(rows, cols , n_proposals))
Beta_Z_z_matrix <- apply(Beta_Z_z_array, 1:2, mean, na.rm = TRUE)

toc()

Beta_Z_z_list[[month_no]] <- flip(terra::setValues(naive_preds[[month_no]], Beta_Z_z_matrix))
# plot(naive_ud_cropped[[i]])
# plot(Beta_Z_z_list[[month_no]])

# exp(beta * Z(x)) * exp(beta * Z(z)) * psi(.) dz
u_x_unnorm <- naive_preds[[month_no]] * Beta_Z_z_list[[month_no]]
# plot(u_x_unnorm)
u_x <- u_x_unnorm / as.numeric(terra::global(u_x_unnorm, fun = "sum", na.rm = TRUE))
names(u_x) <- paste0("MCBM_month_", month_no)
plot(u_x)

MCBM_preds_list[[month_no]] <- u_x

}

```

Plotting

```{r}

for(month_no in 1:12) {

# exp(beta * Z(x))
plot(naive_pred_stack[[month_no]])
plot(naive_pred_stack[[month_no]])
points(as.numeric(buffalo_CLR_year %>% 
                    filter(y == 1 & month == month_no) %>% 
                    dplyr::select(x1) %>% unlist()) - xmin, 
       as.numeric(buffalo_CLR_year %>% 
                    filter(y == 1 & month == month_no) %>% 
                    dplyr::select(y1) %>% unlist()) - ymin)

# exp(beta * Z(z)) * psi(.) dz
plot(Beta_Z_z_list[[month_no]])

# exp(beta * Z(x)) * exp(beta * Z(z)) * psi(.) dz
plot(MCBM_preds_list[[month_no]])
plot(MCBM_preds_list[[month_no]])
points(as.numeric(buffalo_CLR_year %>%
                    filter(y == 1 & month == month_no) %>%
                    dplyr::select(x1) %>% unlist()) - xmin,
       as.numeric(buffalo_CLR_year %>%
                    filter(y == 1 & month == month_no) %>%
                    dplyr::select(y1) %>% unlist()) - ymin)

}

```


Boyce index 

```{r}

boyce_naive_list <- vector(mode = "list", length = 12)
boyce_MCBM_list <- vector(mode = "list", length = 12)

for(month_no in 1:12) {

buffalo_obs <- buffalo_CLR_year_harmonics %>% 
  filter(y == 1 & month == month_no) %>% 
  transmute(x = x1 - xmin, y = y1 - ymin)

naive_raster <- raster(naive_norm_stack[[month_no]])
# boyce_naive_list[[month_no]] <- ecospat.boyce2(naive_raster, buffalo_obs, method = "spearman")
boyce_naive_list[[month_no]] <- ecospat.boyce2(naive_raster, buffalo_obs, method = "pearson")

# }

MCBM_raster <- raster(MCBM_preds_list[[month_no]])
# boyce_MCBM_list[[month_no]] <- ecospat.boyce2(MCBM_raster, buffalo_obs, method = "spearman")
boyce_MCBM_list[[month_no]] <- ecospat.boyce2(MCBM_raster, buffalo_obs, method = "pearson")

}

```


```{r}

spearman_cor_naive <- c()
spearman_cor_MCBM <- c()

for(i in 1:12) {
  
  print(boyce_naive_list[[i]]$cor)
  spearman_cor_naive[i] <- boyce_naive_list[[i]]$cor

  print(boyce_MCBM_list[[i]]$cor)
  spearman_cor_MCBM[i] <- boyce_MCBM_list[[i]]$cor
  
}

mean(spearman_cor_naive)
sd(spearman_cor_naive)

mean(spearman_cor_MCBM)
sd(spearman_cor_MCBM)

```


```{r}

# full_model_2harms_ndvi2_canopy2_herby_naive_cor <- spearman_cor_naive
# mean(full_model_2harms_ndvi2_canopy2_herby_naive_cor)
# sd(full_model_2harms_ndvi2_canopy2_herby_naive_cor)
# 
# full_model_2harms_ndvi2_canopy2_herby_MCBM_cor <- spearman_cor_MCBM
# mean(full_model_2harms_ndvi2_canopy2_herby_MCBM_cor)
# sd(full_model_2harms_ndvi2_canopy2_herby_MCBM_cor)



full_model_2harms_ndvi2_canopy2_herby_elev2_naive_cor <- spearman_cor_naive
mean(full_model_2harms_ndvi2_canopy2_herby_elev2_naive_cor)
sd(full_model_2harms_ndvi2_canopy2_herby_elev2_naive_cor)

full_model_2harms_ndvi2_canopy2_herby_elev2_MCBM_cor <- spearman_cor_MCBM
mean(full_model_2harms_ndvi2_canopy2_herby_elev2_MCBM_cor)
sd(full_model_2harms_ndvi2_canopy2_herby_elev2_MCBM_cor)

```

```{r}

# save.image("workspace images/full_model_2harms_ndvi2_canopy2_herby.RData")
save.image("workspace images/full_model_2harms_ndvi2_canopy2_herby_elev2.RData")

# beep(sound = 2)

```

