---
title: "SSF CLR fitting and prediction"
author: "Scott Forrest"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Import packages

```{r message = FALSE}

options(scipen=999)

# install.packages("forecast")
# install.packages("TwoStepCLogit")

library(tidyverse)

packages <- c("amt", "lubridate", "mgcv", "survival", "terra", "raster", "tictoc", 
              "RColorBrewer", "patchwork", "ecospat", "ggmap", "basemaps", "mapedit", 
              "sf", "ggpubr", "circular", "forecast", "TwoStepCLogit")

walk(packages, require, character.only = T)

```

Without the harmonic terms to have a comparison against the daily harmonic models

```{r two harmonics}

formula_twostep <- y ~ 
  
  ndvi_scaled +
  ndvi_2 +
  canopy_scaled +
  canopy_2 +
  herby_scaled +
  sl +
  log_sl +
  cos_ta +
  
  strata(step_id) +
  cluster(id)

```

Fitting the model

```{r twostep model fit}

month_no <- 3
buffalo_model_data <- buffalo_CLR_year_harmonics %>% dplyr::filter(month == month_no)

months_wet <- c(1:4, 11, 12)
buffalo_model_data_wet <- buffalo_CLR_year_harmonics %>% dplyr::filter(month %in% months_wet)
buffalo_model_data_dry <- buffalo_CLR_year_harmonics %>% dplyr::filter(!month %in% months_wet)

tic()
model_twostep <- Ts.estim(formula = formula_twostep,
         data = buffalo_model_data_wet,
         all.m.1 = TRUE,
         D = "UN(1)",
         itermax = 10000) # default = 2000
toc()

model_twostep_linear_quads_wet <- model_twostep

```

Summary of model fit

```{r two step summary}

model_twostep
# summary(model_twostep)
# print(model_twostep)

```

Reconstructing the Gamma and von Mises distributions from the tentative distributions (from Fieberg et al 2021: Appendix C)

```{r}

# for individuals
# movement_parameters <- read_csv("outputs/buffalo_parametric_indv_movement_params_20230208.csv")
# movement_parameters_year <- movement_parameters %>% filter(id %in% buffalo_year_ids)

# for the population
# gamma$params$shape # 0.438167
# gamma$params$scale # 534.3507
# vonmises$params$kappa # 0.1848126

tentative_shape <- 0.438167
tentative_scale <- 534.3507
tentative_kappa <- 0.1848126

updated_shape <- tentative_shape + model_twostep$beta[which(names(model_twostep$beta) == "log_sl")]
updated_scale <- 1/((1/tentative_scale) - model_twostep$beta[which(names(model_twostep$beta) == "sl")])
updated_kappa <- tentative_kappa + model_twostep$beta[which(names(model_twostep$beta) == "cos_ta")]

```



```{r}

coefs <- unique(hourly_coefs_2harms_quads_long$coef)
# coef_titles <- c("NDVI", "Slope", "Herbaceous vegetation", "Canopy cover", "Step length", "log(Step length)", "cos(Turning angle)")

ggplot() +
    geom_path(data = hourly_coefs_2harms_quads_long %>%
                filter(coef %in% unique(hourly_coefs_2harms_quads_long$coef)[1:4]),
              aes(x = hour, y = value, colour = coef)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    scale_y_continuous(expression(Time-varying~parameter~values~beta)) +
    scale_x_continuous("Hour of the day") +
    scale_color_discrete("Estimate") +
    # ggtitle(coef_titles[i]) +
    theme_classic() +
    theme(legend.position = "bottom")

# ggsave(paste("outputs/plots/clr_fitting/clr_3harmonics_all_quads_20230519.png", sep = ""),
#   width=150, height=90, units="mm", dpi = 300)
# ggsave(paste("outputs/plots/clr_fitting/clr_harmonic_all_pres_20230208.png", sep = ""),
#   width=300, height=180, units="mm", dpi = 300)

ggplot() +
    geom_path(data = hourly_coefs_2harms_quads_long %>% 
                filter(coef %in% unique(hourly_coefs_2harms_quads_long$coef)[c(8,10)]), 
              aes(x = hour, y = value, colour = coef)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    scale_y_continuous(expression(beta)) +
    # scale_x_continuous(breaks = 0:24) +
    scale_color_discrete("Estimate") +
    # ggtitle(coef_titles[i]) +
    theme_classic() +
    theme(legend.position = "bottom")

ggplot() +
    geom_path(data = hourly_coefs_2harms_quads_long %>% 
                filter(coef %in% unique(hourly_coefs_2harms_quads_long$coef)[9]), 
              aes(x = hour, y = value, colour = coef)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    scale_y_continuous(expression(beta)) +
    # scale_x_continuous(breaks = 0:24) +
    scale_color_discrete("Estimate") +
    # ggtitle(coef_titles[i]) +
    theme_classic() +
    theme(legend.position = "bottom")

```




```{r}

x <- seq(0, 1000, length = 500)

# with tentative parameters
plot(x, dgamma(x, shape = tentative_shape, scale = tentative_scale), type = "l")

temporal_movement_params <- hourly_coefs_2harms_quads %>% dplyr::select(c("hour", "shape", "scale", "kappa"))
# write_csv(temporal_movement_params, "outputs/temporal_movement_params_CLR_20230210.csv")

ggplot() +
  stat_function(fun = dgamma, args = list(shape = tentative_shape, scale = tentative_scale)) +
  scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,100)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_classic()

# subset <- temporal_movement_params %>% filter(month < 1)

# ggplot(subset) +
#   stat_function(fun = dgamma, args = list(shape = shape, scale = scale)) +
#   scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,100)) +
#   geom_hline(yintercept = 0, linetype = "dashed") +
#   theme_classic()

```


# Generating predictions

Use the habitat layer script to import habitat layers

# Predictions and cross-validation

Boyce index function

```{r}

ecospat.boyce2 <- function (fit, obs, nclass = 0, window.w = "default", res = 100, 
  PEplot = TRUE, rm.duplicate = TRUE, method = "spearman") 
{
  boycei <- function(interval, obs, fit) {
    pi <- sum(as.numeric(obs >= interval[1] & obs <= interval[2]))/length(obs)
    ei <- sum(as.numeric(fit >= interval[1] & fit <= interval[2]))/length(fit)
    return(round(pi/ei, 10))
  }
  if (inherits(fit, "RasterLayer")) {
    if (is.data.frame(obs) || is.matrix(obs)) {
      obs <- raster::extract(fit, obs)
    }
    fit <- getValues(fit)
    fit <- fit[!is.na(fit)]
    obs <- obs[!is.na(obs)]
  }
  mini <- min(fit, obs)
  maxi <- max(fit, obs)
  if (length(nclass) == 1) {
    if (nclass == 0) {
      if (window.w == "default") {
        window.w <- (max(fit) - min(fit))/10
      }
      vec.mov <- seq(from = mini, to = maxi - window.w, 
        by = (maxi - mini - window.w)/res)
      vec.mov[res + 1] <- vec.mov[res + 1] + 1
      interval <- cbind(vec.mov, vec.mov + window.w)
    }
    else {
      vec.mov <- seq(from = mini, to = maxi, by = (maxi - 
        mini)/nclass)
      interval <- cbind(vec.mov, c(vec.mov[-1], maxi))
    }
  }
  else {
    vec.mov <- c(mini, sort(nclass[!nclass > maxi | nclass < 
      mini]))
    interval <- cbind(vec.mov, c(vec.mov[-1], maxi))
  }
  f <- apply(interval, 1, boycei, obs, fit)
  to.keep <- which(f != "NaN")
  f <- f[to.keep]
  if (length(f) < 2) {
    b <- NA
  }
  else {
    r <- 1:length(f)
    if (rm.duplicate == TRUE) {
      r <- c(1:length(f))[f != c(f[-1], TRUE)]
    }
    b <- cor(f[r], vec.mov[to.keep][r], method = method)
  }
  HS <- apply(interval, 1, sum)/2
  if (length(nclass) == 1 & nclass == 0) {
    HS[length(HS)] <- HS[length(HS)] - 1
  }
  HS <- HS[to.keep]
  if (PEplot == TRUE) {
    plot(HS, f, xlab = "Habitat suitability", ylab = "Predicted/Expected ratio", 
      col = "grey", cex = 0.75)
    points(HS[r], f[r], pch = 19, cex = 0.75)
  }
  return(list(F.ratio = f, cor = round(b, 3), HS = HS))
}

```

Naive approach

```{r}

# naive_pred_stack <- c()
# boyce_naive_list <- vector(mode = "list", length = 24)

# for(i in 1:24) {
  
  # i = 24

# resources <- c(ndvi_stack_scaled[[12 + month_no]], 
#                canopy_scaled,
#                herby_scaled)

resources <- c(ndvi_mean_wet,
               canopy_scaled_cropped,
               herby_scaled_cropped)

# ndvi
ndvi_lin <- resources[[1]]
ndvi_lin <- ndvi_lin * model_twostep$beta[which(names(model_twostep$beta) == "ndvi_scaled")]
plot(ndvi_lin)

ndvi_quad <- resources[[1]]
ndvi_quad <- (ndvi_quad ^ 2) * model_twostep$beta[which(names(model_twostep$beta) == "ndvi_2")]
plot(ndvi_quad)

ndvi_pred <- ndvi_lin  + ndvi_quad
plot(ndvi_pred)
plot(resources[[1]])

# canopy cover 
canopy_lin <- resources[[2]]
canopy_lin <- canopy_lin * model_twostep$beta[which(names(model_twostep$beta) == "canopy_scaled")]
plot(canopy_lin)

canopy_quad <- resources[[2]]
canopy_quad <- (canopy_quad ^ 2) * model_twostep$beta[which(names(model_twostep$beta) == "canopy_2")]
plot(canopy_quad)

canopy_pred <- canopy_lin + canopy_quad
plot(canopy_pred)
plot(resources[[2]])

# herbaceous vegetation
herby_lin <- resources[[3]]
herby_lin <- herby_lin * model_twostep$beta[which(names(model_twostep$beta) == "herby_scaled")]
plot(herby_lin)
plot(resources[[3]])

# combining
naive_pred <- exp(ndvi_pred + canopy_pred + herby_lin) / global(exp(ndvi_pred + canopy_pred + herby_lin), fun = "sum", na.rm = TRUE)[[1]]
plot(naive_pred)
plot(naive_pred)
points(as.numeric(buffalo_CLR_year %>% 
                    filter(y == 1 & month == month_no) %>% 
                    dplyr::select(x1) %>% unlist()), 
       as.numeric(buffalo_CLR_year %>% 
                    filter(y == 1 & month == month_no) %>% 
                    dplyr::select(y1) %>% unlist()))

# naive_pred_stack <- c(naive_pred_stack, naive_pred)

# setting up the buffalo observation data for the relevant month

# buffalo_obs <- buffalo_CLR_year %>% 
#   filter(y == 1 & month == 1 & x1 > xmin & x1 < xmax & y1 > ymin & y1 < ymax) %>% 
#   transmute(x = x1 - xmin, y = y1 - ymin)

buffalo_obs <- buffalo_CLR_year_harmonics %>% 
  filter(y == 1 &  month %in% months_wet) %>% 
  transmute(x = x1, y = y1)

naive_raster <- raster(naive_pred)
# naive_raster <- raster(naive_pred_cropped)

# plot(naive_pred)
# plot(naive_raster)

# plot(naive_cropped_raster[[1]])

# boyce_naive <- ecospat.boyce2(naive_cropped_raster[[1]], buffalo_obs,
#                               # method = "spearman"
#                               method = "pearson",
#                               # method = "kendall"
#                               )

boyce_naive_allhours <- ecospat.boyce2(naive_raster, buffalo_obs,
                              method = "spearman"
                              # method = "pearson",
                              # method = "kendall"
                              )

boyce_naive_allhours

# }

# naive_raster_2harms <- naive_raster
# boyce_naive_list_2harms <- boyce_naive_list

# naive_raster_daily_2harms_quads <- naive_raster
# boyce_naive_raster_daily_2harms_quads <- boyce_naive_list

```


```{r}

naive_pred_linear_df <- as.data.frame(naive_pred, xy = TRUE)

# smaller 100 x 100 extent
xmin <- 27500
xmax <- 35000
ymin <- -1438500
ymax <- -1431000

crop_extent <- ext(xmin, xmax, ymin, ymax)
naive_pred_cropped <- terra::crop(naive_pred, crop_extent)
# ext(naive_ud_cropped) <- c(xmin - xmin, 
#                             xmax - xmin, 
#                             ymin - ymin, 
#                             ymax - ymin)


naive_pred_linear_cropped_df <- as.data.frame(naive_pred_cropped, xy = TRUE)

for(i in 1:24) {
  
plot <- ggplot() +
  geom_raster(data = naive_pred_linear_cropped_df, aes(x = x, y = y, fill = mean)) +
  geom_point(data = buffalo_CLR_year_harmonics %>% 
               filter(y == 1 & month %in% months_wet & hour == ifelse(i == 24, 0, i)), 
             aes(x = x1, y = y1), colour = "red", size = 0.5) +
  coord_equal() +
  scale_x_continuous(limits = c(xmin, xmax )) + # - xmin
  scale_y_continuous(limits = c(ymin, ymax )) + # - ymin
  scale_fill_viridis_c("Habitat suitability", option = "D", direction = 1) +
  # , limits = c(0.00000001185235, 0.0000006519935)
  # scale_fill_distiller("Habitat suitability", palette = "Reds", direction = 1) +
  ggtitle(paste0("Hour ", i)) +
  theme_classic() +
  theme(legend.position = "right")

print(plot)

ggsave(paste0("outputs/plots/hourly_preds/hourly_preds_allhours_", i, "_20230525.png"), width=150, height=90, units="mm", dpi = 300)

}

```



```{r}

# for(i in 1:24) {
#   plot(naive_pred_stack[[i]])
# }
# 
# for(i in 1:24) {
#   # plot(naive_pred_stack[[i]])
#   plot(naive_pred_stack[[i]])
#   points(as.numeric(buffalo_CLR_year %>% filter(y == 1 & month == month_no & hour == ifelse(i == 24, 0, i)) %>% dplyr::select(x1) %>% unlist()), 
#        as.numeric(buffalo_CLR_year %>% filter(y == 1 & month == month_no & hour == ifelse(i == 24, 0, i)) %>% dplyr::select(y1) %>% unlist()))
# }

```


```{r}

# spearman_cor <- c()
# 
# for(i in 1:24) {
#   print(boyce_naive_list[[i]]$cor)
#   spearman_cor[i] <- boyce_naive_list[[i]]$cor
# }
# 
# mean(spearman_cor)
# sd(spearman_cor)

```

Subsetting the spatial extent to predict over

```{r}

# finding the minimum and maximum extent (using the pseudo-absences as a buffer around the points)

buffalo_CLR_year_harmonics %>% dplyr::summarise(xmin = min(x2), xmax = max(x2),
                                                ymin = min(y2), ymax = max(y2))

buffalo_CLR_year_harmonics %>% dplyr::summarise(xmin = min(x1), xmax = max(x1),
                                                ymin = min(y1), ymax = max(y1))

subset_extent <- terra::ext(round(min(buffalo_CLR_year_harmonics$x2), digits = - 2), round(max(buffalo_CLR_year_harmonics$x2), digits = - 2),
           round(min(buffalo_CLR_year_harmonics$y2), digits = - 2), round(max(buffalo_CLR_year_harmonics$y2), digits = - 2))

resources_cropped <- terra::crop(resources, subset_extent)
plot(resources_cropped)
plot(resources_cropped[[1]])
points(as.numeric(buffalo_CLR_year %>% 
                    filter(y == 1) %>% 
                    dplyr::select(x1) %>% unlist()), 
       as.numeric(buffalo_CLR_year %>% 
                    filter(y == 1) %>% 
                    dplyr::select(y1) %>% unlist()))

```

Naive approach

```{r}

# naive_pred_stack <- c()
# boyce_naive_list <- vector(mode = "list", length = 24)

# for(i in 1:24) {

# ndvi
ndvi_lin <- resources_cropped[[1]]
ndvi_lin <- ndvi_lin * model_twostep$beta[which(names(model_twostep$beta) == "ndvi_scaled")]
plot(ndvi_lin)

ndvi_quad <- resources_cropped[[1]]
ndvi_quad <- (ndvi_quad ^ 2) * model_twostep$beta[which(names(model_twostep$beta) == "ndvi_2")]
plot(ndvi_quad)

ndvi_pred <- ndvi_lin  + ndvi_quad
plot(ndvi_pred)
plot(resources_cropped[[1]])

# canopy cover 
canopy_lin <- resources_cropped[[2]]
canopy_lin <- canopy_lin * model_twostep$beta[which(names(model_twostep$beta) == "canopy_scaled")]
plot(canopy_lin)

canopy_quad <- resources_cropped[[2]]
canopy_quad <- (canopy_quad ^ 2) * model_twostep$beta[which(names(model_twostep$beta) == "canopy_2")]
plot(canopy_quad)

canopy_pred <- canopy_lin + canopy_quad
plot(canopy_pred)
plot(resources_cropped[[2]])

# herbaceous vegetation
herby_lin <- resources_cropped[[3]]
herby_lin <- herby_lin * model_twostep$beta[which(names(model_twostep$beta) == "herby_scaled")]
plot(herby_lin)
plot(resources_cropped[[3]])

# combining
naive_pred <- exp(ndvi_pred + canopy_pred + herby_lin) / global(exp(ndvi_pred + canopy_pred + herby_lin), fun = "sum", na.rm = TRUE)[[1]]
plot(naive_pred)
plot(naive_pred)
points(as.numeric(buffalo_CLR_year %>% 
                    filter(y == 1 & month == month_no) %>% 
                    dplyr::select(x1) %>% unlist()), 
       as.numeric(buffalo_CLR_year %>% 
                    filter(y == 1 & month == month_no) %>% 
                    dplyr::select(y1) %>% unlist()))

# naive_pred_stack <- c(naive_pred_stack, naive_pred)

# setting up the buffalo observation data for the relevant month

# buffalo_obs <- buffalo_CLR_year %>% 
#   filter(y == 1 & month == 1 & x1 > xmin & x1 < xmax & y1 > ymin & y1 < ymax) %>% 
#   transmute(x = x1 - xmin, y = y1 - ymin)

buffalo_obs <- buffalo_CLR_year %>% 
  filter(y == 1 & month == month_no) %>% 
  transmute(x = x1, y = y1)

naive_raster <- raster(naive_pred)
# naive_raster <- raster(naive_pred_cropped)

# plot(naive_pred)
# plot(naive_raster)

# plot(naive_cropped_raster[[1]])

# boyce_naive <- ecospat.boyce2(naive_cropped_raster[[1]], buffalo_obs,
#                               # method = "spearman"
#                               method = "pearson",
#                               # method = "kendall"
#                               )

boyce_naive <- ecospat.boyce2(naive_raster, buffalo_obs,
                              method = "spearman"
                              # method = "pearson",
                              # method = "kendall"
                              )

boyce_naive

# boyce_naive_list[[i]] <- boyce_naive

# }

# naive_raster_2harms <- naive_raster
# boyce_naive_list_2harms <- boyce_naive_list

# naive_raster_daily_2harms_quads <- naive_raster
# boyce_naive_raster_daily_2harms_quads <- boyce_naive_list

```