---
title: "SSF CLR fitting and prediction"
author: "Scott Forrest"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Import packages

```{r message = FALSE}

options(scipen=999)

# install.packages("forecast")
# install.packages("TwoStepCLogit")

library(tidyverse)

packages <- c("amt", "lubridate", "mgcv", "survival", "terra", "raster", "tictoc", 
              "RColorBrewer", "patchwork", "ecospat", "ggmap", "basemaps", "mapedit", 
              "sf", "ggpubr", "circular", "forecast", "TwoStepCLogit")

walk(packages, require, character.only = T)

```

With 2 pairs of harmonics 

```{r two harmonics}

# auto.arima()

formula_twostep <- y ~ 
  
  ndvi_scaled +
  yday_s1:ndvi_scaled +
  yday_s2:ndvi_scaled +
  # yday_s3:ndvi_scaled +
  # yday_s4:ndvi_scaled +
  # yday_s5:ndvi_scaled +
  # yday_s6:ndvi_scaled +
  
  yday_c1:ndvi_scaled +
  yday_c2:ndvi_scaled +
  # yday_c3:ndvi_scaled +
  # yday_c4:ndvi_scaled +
  # yday_c5:ndvi_scaled +
  # yday_c6:ndvi_scaled +
  
  ndvi_2 +
  yday_s1:ndvi_2 +
  yday_s2:ndvi_2 +
  # yday_s3:ndvi_2 +
  # yday_s4:ndvi_2 +
  # yday_s5:ndvi_2 +
  # yday_s6:ndvi_2 +
  
  yday_c1:ndvi_2 +
  yday_c2:ndvi_2 +
  # yday_c3:ndvi_2 +
  # yday_c4:ndvi_2 +
  # yday_c5:ndvi_2 +
  # yday_c6:ndvi_2 +

  canopy_scaled +
  yday_s1:canopy_scaled +
  yday_s2:canopy_scaled +
  # yday_s3:canopy_scaled +
  # yday_s4:canopy_scaled +
  # yday_s5:canopy_scaled +
  # yday_s6:canopy_scaled +
  
  yday_c1:canopy_scaled +
  yday_c2:canopy_scaled +
  # yday_c3:canopy_scaled +
  # yday_c4:canopy_scaled +
  # yday_c5:canopy_scaled +
  # yday_c6:canopy_scaled +
  
  canopy_2 +
  yday_s1:canopy_2 +
  yday_s2:canopy_2 +
  # yday_s3:canopy_2 +
  # yday_s4:canopy_2 +
  # yday_s5:canopy_2 +
  # yday_s6:canopy_2 +
  
  yday_c1:canopy_2 +
  yday_c2:canopy_2 +
  # yday_c3:canopy_2 +
  # yday_c4:canopy_2 +
  # yday_c5:canopy_2 +
  # yday_c6:canopy_2 +
  
  sl +
  yday_s1:sl +
  yday_s2:sl +
  # yday_s3:sl +
  # yday_s4:sl +
  # yday_s5:sl +
  # yday_s6:sl +
  
  yday_c1:sl +
  yday_c2:sl +
  # yday_c3:sl +
  # yday_c4:sl +
  # yday_c5:sl +
  # yday_c6:sl +
  
  log_sl +
  yday_s1:log_sl +
  yday_s2:log_sl +
  # yday_s3:log_sl +
  # yday_s4:log_sl +
  # yday_s5:log_sl +
  # yday_s6:log_sl +
  
  yday_c1:log_sl +
  yday_c2:log_sl +
  # yday_c3:log_sl +
  # yday_c4:log_sl +
  # yday_c5:log_sl +
  # yday_c6:log_sl +
  
  cos_ta +
  yday_s1:cos_ta +
  yday_s2:cos_ta +
  # yday_s3:cos_ta +
  # yday_s4:cos_ta +
  # yday_s5:cos_ta +
  # yday_s6:cos_ta +
  
  yday_c1:cos_ta +
  yday_c2:cos_ta +
  # yday_c3:cos_ta +
  # yday_c4:cos_ta +
  # yday_c5:cos_ta +
  # yday_c6:cos_ta +
  
  strata(step_id) +
  cluster(id)

```


With 3 pairs of harmonics 

```{r three harmonics}

# auto.arima()

formula_twostep <- y ~ 
  
  ndvi_scaled +
  yday_s1:ndvi_scaled +
  yday_s2:ndvi_scaled +
  yday_s3:ndvi_scaled +
  # yday_s4:ndvi_scaled +
  # yday_s5:ndvi_scaled +
  # yday_s6:ndvi_scaled +
  
  yday_c1:ndvi_scaled +
  yday_c2:ndvi_scaled +
  yday_c3:ndvi_scaled +
  # yday_c4:ndvi_scaled +
  # yday_c5:ndvi_scaled +
  # yday_c6:ndvi_scaled +
  
  ndvi_2 +
  yday_s1:ndvi_2 +
  yday_s2:ndvi_2 +
  yday_s3:ndvi_2 +
  # yday_s4:ndvi_2 +
  # yday_s5:ndvi_2 +
  # yday_s6:ndvi_2 +
  
  yday_c1:ndvi_2 +
  yday_c2:ndvi_2 +
  yday_c3:ndvi_2 +
  # yday_c4:ndvi_2 +
  # yday_c5:ndvi_2 +
  # yday_c6:ndvi_2 +

  canopy_scaled +
  yday_s1:canopy_scaled +
  yday_s2:canopy_scaled +
  yday_s3:canopy_scaled +
  # yday_s4:canopy_scaled +
  # yday_s5:canopy_scaled +
  # yday_s6:canopy_scaled +
  
  yday_c1:canopy_scaled +
  yday_c2:canopy_scaled +
  yday_c3:canopy_scaled +
  # yday_c4:canopy_scaled +
  # yday_c5:canopy_scaled +
  # yday_c6:canopy_scaled +
  
  canopy_2 +
  yday_s1:canopy_2 +
  yday_s2:canopy_2 +
  yday_s3:canopy_2 +
  # yday_s4:canopy_2 +
  # yday_s5:canopy_2 +
  # yday_s6:canopy_2 +
  
  yday_c1:canopy_2 +
  yday_c2:canopy_2 +
  yday_c3:canopy_2 +
  # yday_c4:canopy_2 +
  # yday_c5:canopy_2 +
  # yday_c6:canopy_2 +
  
  sl +
  yday_s1:sl +
  yday_s2:sl +
  yday_s3:sl +
  # yday_s4:sl +
  # yday_s5:sl +
  # yday_s6:sl +
  
  yday_c1:sl +
  yday_c2:sl +
  yday_c3:sl +
  # yday_c4:sl +
  # yday_c5:sl +
  # yday_c6:sl +
  
  log_sl +
  yday_s1:log_sl +
  yday_s2:log_sl +
  yday_s3:log_sl +
  # yday_s4:log_sl +
  # yday_s5:log_sl +
  # yday_s6:log_sl +
  
  yday_c1:log_sl +
  yday_c2:log_sl +
  yday_c3:log_sl +
  # yday_c4:log_sl +
  # yday_c5:log_sl +
  # yday_c6:log_sl +
  
  cos_ta +
  yday_s1:cos_ta +
  yday_s2:cos_ta +
  yday_s3:cos_ta +
  # yday_s4:cos_ta +
  # yday_s5:cos_ta +
  # yday_s6:cos_ta +
  
  yday_c1:cos_ta +
  yday_c2:cos_ta +
  yday_c3:cos_ta +
  # yday_c4:cos_ta +
  # yday_c5:cos_ta +
  # yday_c6:cos_ta +
  
  strata(step_id) +
  cluster(id)

```


```{r}

# formula_twostep <- y ~ 
#   
#   poly(ndvi_scaled, degree = 2, raw = TRUE) +
#   poly(ndvi_s1, degree = 2, raw = TRUE) +
#   poly(ndvi_s2, degree = 2, raw = TRUE) +
#   poly(ndvi_c1, degree = 2, raw = TRUE) +
#   poly(ndvi_c2, degree = 2, raw = TRUE) +
#   
#   poly(canopy_scaled, degree = 2, raw = TRUE) +
#   poly(canopy_s1, degree = 2, raw = TRUE) +
#   poly(canopy_s2, degree = 2, raw = TRUE) +
#   poly(canopy_c1, degree = 2, raw = TRUE) +
#   poly(canopy_c2, degree = 2, raw = TRUE) +
#   
#   sl +
#   sl_s1 +
#   sl_s2 +
#   sl_c1 +
#   sl_c2 +
#   
#   log_sl +
#   log_sl_s1 +
#   log_sl_s2 +
#   log_sl_c1 +
#   log_sl_c2 +
#   
#   cos_ta +
#   cos_ta_s1 +
#   cos_ta_s2 +
#   cos_ta_c1 +
#   cos_ta_c2 +
#   
#   strata(step_id) +
#   cluster(id)

```


Fitting the model

```{r twostep model fit}



tic()
model_twostep <- Ts.estim(formula = formula_twostep,
         data = buffalo_CLR_year_harmonics,
         # random = , # default = all covariates
         all.m.1 = TRUE,
         D = "UN(1)",
         itermax = 10000) # default = 2000
toc()

# model_twostep_2harms <- model_twostep
model_twostep <- model_twostep_3harms

```


Summary of model fit

```{r two step summary}

model_twostep
# summary(model_twostep)
# print(model_twostep)

```




```{r}

# 1:length(model_twostep$beta)
# names(model_twostep$beta)
# as.numeric(model_twostep$beta)
# as.numeric(model_twostep$se)
# as.numeric(diag(model_twostep$D))

coefs_df_function <- function(clogit_model) {
  
  return(data.frame("index" = 1:length(model_twostep$beta), 
                    "coef" = names(model_twostep$beta), 
                    "value" = as.numeric(model_twostep$beta), 
                    "se" = as.numeric(model_twostep$se),
                    "ind.var" = as.numeric(diag(model_twostep$D))))
}

clogit_model_coefs <- coefs_df_function(model_twostep)

# store the coefficients of each model fit
# clogit_model_coefs_2harms <- clogit_model_coefs
clogit_model_coefs_3harms <- clogit_model_coefs

# clogit_model_coefs <- clogit_model_coefs_2harms
clogit_model_coefs <- clogit_model_coefs_3harms

```

Calculating the coefficient curves across the year (by yday)

```{r temporal coefficients 2 harmonics}

# clogit_coef_df <- clogit_model_coefs

x_yday <- seq(1, 365, 1)

y_yday <- function(x,
                    lin_coef,
                    sin_coef1,
                    sin_coef2,
                    cos_coef1,
                    cos_coef2) {

  lin_coef +
    (sin_coef1 * sin(2 * pi * x / 365)) +
    (sin_coef2 * sin(4 * pi * x / 365)) +
    (cos_coef1 * cos(2 * pi * x / 365)) +
    (cos_coef2 * cos(4 * pi * x / 365))
}



harmonic_coefs_function <- function(clogit_coef_df) {

y_ndvi <- map_dbl(x_yday,
             y_yday,
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_scaled")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_scaled:yday_s1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_scaled:yday_s2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_scaled:yday_c1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_scaled:yday_c2")])

y_ndvi_quad <- map_dbl(x_yday,
             y_yday,
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_2:yday_s1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_2:yday_s2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_2:yday_c1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_2:yday_c2")])

y_canopy <- map_dbl(x_yday,
             y_yday,
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_scaled")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_scaled:yday_s1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_scaled:yday_s2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_scaled:yday_c1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_scaled:yday_c2")])

y_canopy_quad <- map_dbl(x_yday,
             y_yday,
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_2:yday_s1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_2:yday_s2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_2:yday_c1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_2:yday_c2")])

y_sl <- map_dbl(x_yday,
             y_yday,
             clogit_coef_df$value[which(clogit_coef_df$coef == "sl")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "sl:yday_s1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "sl:yday_s2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "sl:yday_c1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "sl:yday_c2")])

y_log_sl <- map_dbl(x_yday,
             y_yday,
             clogit_coef_df$value[which(clogit_coef_df$coef == "log_sl")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "log_sl:yday_s1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "log_sl:yday_s2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "log_sl:yday_c1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "log_sl:yday_c2")])

y_cos_ta <- map_dbl(x_yday,
             y_yday,
             clogit_coef_df$value[which(clogit_coef_df$coef == "cos_ta")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "cos_ta:yday_s1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "cos_ta:yday_s2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "cos_ta:yday_c1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "cos_ta:yday_c2")])

return(data.frame("yday" = x_yday,
                  "NDVI" = y_ndvi, 
                  "NDVI_quadratic" = y_ndvi_quad, 
                  "Canopy_cover" = y_canopy, 
                  "Canopy_cover_quadratic" = y_canopy_quad, 
                  "Step_length" = y_sl,
                  "log_Step_length" = y_log_sl, 
                  "cos_Turning_angle" = y_cos_ta))

}


# clogit_model_coefs$value[which(clogit_model_coefs$coef == "cos_ta:yday_s1")]

# running the function
harmonic_coefs_df <- harmonic_coefs_function(clogit_model_coefs)

# harmonic_coefs_df_2harms <- harmonic_coefs_df

```

For the 3 harmonic model

```{r temporal coefficients 3 harmonics}

# clogit_coef_df <- clogit_model_coefs

x_yday <- seq(1, 365, 1)

y_yday <- function(x,
                    lin_coef,
                    sin_coef1,
                    sin_coef2,
                    sin_coef3,
                    cos_coef1,
                    cos_coef2,
                    cos_coef3) {

  lin_coef +
    (sin_coef1 * sin(2 * pi * x / 365)) +
    (sin_coef2 * sin(4 * pi * x / 365)) +
    (sin_coef3 * sin(6 * pi * x / 365)) +
    (cos_coef1 * cos(2 * pi * x / 365)) +
    (cos_coef2 * cos(4 * pi * x / 365)) +
    (cos_coef3 * cos(6 * pi * x / 365))
}



harmonic_coefs_function <- function(clogit_coef_df) {

y_ndvi <- map_dbl(x_yday,
             y_yday,
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_scaled")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_scaled:yday_s1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_scaled:yday_s2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_scaled:yday_s3")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_scaled:yday_c1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_scaled:yday_c2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_scaled:yday_c3")])

y_ndvi_quad <- map_dbl(x_yday,
             y_yday,
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_2:yday_s1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_2:yday_s2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_2:yday_s3")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_2:yday_c1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_2:yday_c2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "ndvi_2:yday_c3")])

y_canopy <- map_dbl(x_yday,
             y_yday,
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_scaled")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_scaled:yday_s1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_scaled:yday_s2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_scaled:yday_s3")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_scaled:yday_c1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_scaled:yday_c2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_scaled:yday_c3")])

y_canopy_quad <- map_dbl(x_yday,
             y_yday,
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_2:yday_s1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_2:yday_s2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_2:yday_s3")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_2:yday_c1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_2:yday_c2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "canopy_2:yday_c3")])

y_sl <- map_dbl(x_yday,
             y_yday,
             clogit_coef_df$value[which(clogit_coef_df$coef == "sl")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "sl:yday_s1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "sl:yday_s2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "sl:yday_s3")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "sl:yday_c1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "sl:yday_c2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "sl:yday_c3")])

y_log_sl <- map_dbl(x_yday,
             y_yday,
             clogit_coef_df$value[which(clogit_coef_df$coef == "log_sl")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "log_sl:yday_s1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "log_sl:yday_s2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "log_sl:yday_s3")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "log_sl:yday_c1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "log_sl:yday_c2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "log_sl:yday_c3")])

y_cos_ta <- map_dbl(x_yday,
             y_yday,
             clogit_coef_df$value[which(clogit_coef_df$coef == "cos_ta")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "cos_ta:yday_s1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "cos_ta:yday_s2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "cos_ta:yday_s3")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "cos_ta:yday_c1")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "cos_ta:yday_c2")],
             clogit_coef_df$value[which(clogit_coef_df$coef == "cos_ta:yday_c3")])

return(data.frame("yday" = x_yday,
                  "NDVI" = y_ndvi, 
                  "NDVI_quadratic" = y_ndvi_quad, 
                  "Canopy_cover" = y_canopy, 
                  "Canopy_cover_quadratic" = y_canopy_quad, 
                  "Step_length" = y_sl,
                  "log_Step_length" = y_log_sl, 
                  "cos_Turning_angle" = y_cos_ta))

}


# clogit_model_coefs$value[which(clogit_model_coefs$coef == "cos_ta:yday_s1")]

# running the function
harmonic_coefs_df <- harmonic_coefs_function(clogit_model_coefs)

harmonic_coefs_df_3harms <- harmonic_coefs_df

```


Reconstructing the Gamma and von Mises distributions from the tentative distributions (from Fieberg et al 2021: Appendix C)

```{r}

# for individuals
# movement_parameters <- read_csv("outputs/buffalo_parametric_indv_movement_params_20230208.csv")
# movement_parameters_year <- movement_parameters %>% filter(id %in% buffalo_year_ids)

# for the population
# gamma$params$shape # 0.438167
# gamma$params$scale # 534.3507
# vonmises$params$kappa # 0.1848126

tentative_shape <- 0.438167
tentative_scale <- 534.3507
tentative_kappa <- 0.1848126

harmonic_coefs_df <- harmonic_coefs_df %>% mutate(shape = tentative_shape + log_Step_length,
                             scale = 1/((1/tentative_scale) - Step_length),
                             kappa = tentative_kappa + cos_Turning_angle)

# write_csv(harmonic_coefs_df, "outputs/harmonic_coefs_df_with_movement_params_ndvi_canopy_quads_20230519.csv")

# turning into a long data frame
harmonic_coefs <- pivot_longer(harmonic_coefs_df, cols = !1, names_to = "coef")

# write_csv(harmonic_coefs, "outputs/harmonic_coefs_df_long_with_movement_params_ndvi_canopy_quads_20230519.csv")

```


```{r}

x <- seq(0,500,1)

# step length distribution
plot(x, dgamma(x, 
       shape = harmonic_coefs %>% filter(yday == 1) %>% 
         dplyr::filter(coef == "shape") %>% dplyr::pull(value),
       scale = harmonic_coefs %>% filter(yday == 1) %>% 
         dplyr::filter(coef == "scale") %>% dplyr::pull(value)),
     type = "l",
     main = "Step length distribution",
     xlab = "Distance (m)",
     ylab = "Density")

xvm <- seq(-pi, pi, 0.01)

plot(xvm, dvonmises(xvm, 
          mu = 0, 
          kappa = harmonic_coefs %>% filter(yday == 1) %>% 
            dplyr::filter(coef == "kappa") %>% dplyr::pull(value),
          log = FALSE),
     type = "l",
     main = "Turning angle distribution",
     xlab = "Turning angle (rad)",
     ylab = "Density",
     ylim = c(0,0.21))


```



```{r}

n <- 1e5

shape <- harmonic_coefs %>% filter(id == "Weighted mean" & month == 12) %>% dplyr::filter(coef == "shape") %>% dplyr::pull(value)
scale <- harmonic_coefs %>% filter(id == "Weighted mean" & month == 12) %>% dplyr::filter(coef == "scale") %>% dplyr::pull(value)
kappa <- harmonic_coefs %>% filter(id == "Weighted mean" & month == 12) %>% dplyr::filter(coef == "kappa") %>% dplyr::pull(value)

sl <- rgamma(n, shape = shape, scale = scale)
hist(sl, breaks = 100)
ta <- as.numeric(Rfast::rvonmises(n, m = pi, k = 0.78) - pi)
hist(ta, breaks = 100)

x_s <- sl * sin(ta)
y_s <- sl * cos(ta)
xy_mk <- data.frame("x_pos" = x_s, "y_pos" = y_s)

ggplot(data = xy_mk, aes(x = x_pos, y = y_pos)) +
  geom_point(alpha = 0.1) +
  # geom_density_2d_filled(data = xy_mk, aes(x = x_pos, y = y_pos)) +
  # geom_hex(bins = 30) +
  # geom_bin_2d(bins = 50) +
  scale_x_continuous("x position") +
  scale_y_continuous("y position") +
  # scale_fill_viridis_c() +
  geom_point(aes(x = 0, y = 0), colour = "red", alpha = 0.25) +
  coord_equal() +
  theme_bw()

```



```{r}

coefs <- unique(harmonic_coefs$coef)
# coef_titles <- c("NDVI", "Slope", "Herbaceous vegetation", "Canopy cover", "Step length", "log(Step length)", "cos(Turning angle)")

for(i in 1:length(coefs)) {
  
  i = 1
  
  gg <- ggplot() +
  # geom_line(data = harmonic_coefs %>% 
  #             filter(coef %in% unique(harmonic_coefs$coef)), # %>% # [1:4]
  #             # filter(coef == coefs[i]), 
  #           aes(x = yday, y = value, group = id), colour = "grey") + # , colour = factor(id)
    geom_line(data = harmonic_coefs %>% 
                filter(coef == coefs[i]), aes(x = yday, y = value, colour = factor(coef))) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    scale_y_continuous(expression(Time-varying~parameter~values~beta)) +
    scale_x_continuous() + # breaks = seq(0,)
    scale_color_discrete("Estimate") +
    ggtitle(coefs[i]) +
    theme_classic() +
    theme(legend.position = "bottom")
  
  print(gg)

  # ggsave(paste("outputs/plots/clr_fitting/clr_harmonic_20230208_", coef_titles[i], ".png", sep = ""),
  # width=150, height=90, units="mm", dpi = 300)
  
}


# with just the means 

ggplot() +
    geom_path(data = harmonic_coefs %>% 
                # filter(id %in% unique(harmonic_coefs$id)[11]) %>% 
                filter(coef %in% unique(harmonic_coefs$coef)[1:7]), 
              aes(x = yday, y = value, colour = coef)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    scale_y_continuous(expression(Time-varying~parameter~values~beta)) +
    scale_x_continuous("Day of the year") +
    scale_color_discrete("Estimate") +
    # ggtitle(coef_titles[i]) +
    theme_classic() +
    theme(legend.position = "bottom")

ggsave(paste("outputs/plots/clr_fitting/clr_3harmonics_all_quads_20230519.png", sep = ""),
  width=150, height=90, units="mm", dpi = 300)
# ggsave(paste("outputs/plots/clr_fitting/clr_harmonic_all_pres_20230208.png", sep = ""),
#   width=300, height=180, units="mm", dpi = 300)


ggplot() +
    geom_path(data = harmonic_coefs %>% 
                filter(id %in% unique(harmonic_coefs$id)[11]) %>% 
                filter(coef %in% unique(harmonic_coefs$coef)[8:10]), 
              aes(x = month, y = value, colour = coef)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    scale_y_continuous(expression(beta)) +
    scale_x_continuous(breaks = 0:12) +
    scale_color_discrete("Estimate") +
    # ggtitle(coef_titles[i]) +
    theme_classic() +
    theme(legend.position = "bottom")


ggplot() +
    geom_path(data = harmonic_coefs %>% 
                filter(id %in% unique(harmonic_coefs$id)[11]) %>% 
                filter(coef %in% unique(harmonic_coefs$coef)[c(8,10)]), 
              aes(x = month, y = value, colour = coef)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    scale_y_continuous(expression(beta)) +
    scale_x_continuous(breaks = 0:12) +
    scale_color_discrete("Estimate") +
    # ggtitle(coef_titles[i]) +
    theme_classic() +
    theme(legend.position = "bottom")

harmonic_coefs %>% filter(id %in% unique(harmonic_coefs$id)[11])

```




```{r}

x <- seq(0, 1000, length = 500)
dgamma(x, shape = tentative_shape, scale = tentative_scale)

temporal_movement_params <- harmonic_coefs_df %>% filter(id == "Weighted mean") %>% dplyr::select(c("month", "shape", "scale", "kappa"))
# write_csv(temporal_movement_params, "outputs/temporal_movement_params_CLR_20230210.csv")

ggplot() +
  stat_function(fun = dgamma, args = list(shape = tentative_shape, scale = tentative_scale)) +
  scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,100)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_classic()


subset <- temporal_movement_params %>% filter(month < 1)

ggplot(subset) +
  stat_function(fun = dgamma, args = list(shape = shape, scale = scale)) +
  scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,100)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_classic()

```


# Generating predictions

Importing habitat layers

```{r}

ndvi_projected <- rast("mapping/cropped rasters/ndvi_GEE_projected_watermask20230207.tif")

# scaling from the buffalo data
scaling_mean <- as.numeric(attributes(buffalo_CLR_year$ndvi_scaled)[1]) # 0.3125089
scaling_sd <- as.numeric(attributes(buffalo_CLR_year$ndvi_scaled)[2]) # 0.143179 for July 2018

ndvi_stack_scaled <- (ndvi_projected - scaling_mean) / scaling_sd
# plot(ndvi_stack_scaled)
# time_vector <- terra::time(ndvi_projected)
# terra::time(ndvi_stack_scaled) <- NULL
# terra::time(ndvi_stack_scaled) <- lubridate::date(as.POSIXlt(as.POSIXlt("1970/01/01 00:00:00") + time_vector))

# ndvi_above0_stack_scaled <- rast("mapping/cropped rasters/ndvi_above0_stack_scaled_20230210.tif")

```

For plotting with terra 

```{r}

# elev <- rast("mapping/cropped rasters/DEM_H_raster.tif")
slope <- rast("mapping/cropped rasters/slope_raster.tif")
veg_herby <- rast("mapping/cropped rasters/veg_herby.tif")
canopy_cover <- rast("mapping/cropped rasters/canopy_cover.tif")
# 
# ndvi_july2018 <- ndvi_projected[[7]]
# 
# names(slope) <- "Slope"
# names(veg_herby) <- "Herbaceous vegetation"
# names(canopy_cover) <- "Canopy cover"
# names(ndvi_july2018) <- "NDVI July 2018"
# 
# # plot(elev)
# plot(slope)
# plot(veg_herby)
# plot(canopy_cover)
# plot(ndvi_july2018)
# 
# plot_stack <- c(ndvi_july2018, slope, veg_herby, canopy_cover)
# 
# png(filename = "outputs/plots/ndvi_slope_herb_canopy_layers_20230208.png",
#     width = 300, height = 175, units = "mm", res = 300)
# plot(plot_stack)
# dev.off()

```

Scaled rasters

```{r}

# hist(buffalo_CLR_year$slope_scaled)
# slope_scaled <- (slope - attr(buffalo_CLR_year$slope_scaled, "scaled:center")) / attr(buffalo_CLR_year$slope_scaled, "scaled:scale")
# plot(slope_scaled)
# 
# writeRaster(slope_scaled, "mapping/cropped rasters/slope_scaled_by_buffalo_data_20230210.tif")
slope_scaled <- rast("mapping/cropped rasters/slope_scaled_by_buffalo_data_20230210.tif")

# hist(buffalo_CLR_year$canopy_scaled)
# canopy_scaled <- ((canopy_cover / 100) - attr(buffalo_CLR_year$canopy_scaled, "scaled:center")) / attr(buffalo_CLR_year$canopy_scaled, "scaled:scale")
# plot(canopy_scaled)
# 
# writeRaster(canopy_scaled, "mapping/cropped rasters/canopy_scaled_by_buffalo_data_20230210.tif")
canopy_scaled <- rast("mapping/cropped rasters/canopy_scaled_by_buffalo_data_20230210.tif")

```


Checking distributions of covariates vs used and random locations

```{r}

ndvi_values_month1 <- values(ndvi_above0_stack_scaled[[13]])
canopy_values <- values(canopy_cover)
# slope_values <- values(slope_scaled)
# herby_values <- values(veg_herby)

enviro_values <- data.frame("Data" = rep("Environmental", length(values(canopy_cover))), "pres" = rep(1, length(values(canopy_cover))),  "ndvi" = ndvi_values_month1, "canopy" = canopy_values, "slope" = slope_values, "herby" = herby_values)
names(enviro_values) <- c("Data", "pres", "ndvi", "canopy", "slope", "herby")

buffalo_pres_values <- buffalo_CLR_year %>% filter(y == 1 & year == 2019 & month == 3) %>%  transmute(Data = "Presence", pres = y, ndvi = ndvi_temporal, canopy = canopy_cover, slope = slope_end, herby = veg_herby)

buffalo_background_values <- buffalo_CLR_year %>% filter(y == 0 & year == 2019 & month == 3) %>%  transmute(Data = "Random steps", pres = y, ndvi = ndvi_temporal, canopy = canopy_cover, slope = slope_end, herby = veg_herby)

enviro_buffalo_values <- rbind(enviro_values, buffalo_pres_values, buffalo_background_values)

# ggplot(enviro_buffalo_values) +
#   geom_density(aes(x = ndvi, fill = method), alpha = 0.5) +
#   geom_density(data = buffalo_CLR_year %>% filter(y == 1, year == 2019 & month == 1), 
#                aes(x = ndvi_temporal), fill = "orange", alpha = 0.5) +
#   geom_density(data = buffalo_CLR_year %>% filter(y == 0, year == 2019 & month == 1), 
#                aes(x = ndvi_temporal), linetype = "dashed", colour = "red", alpha = 0.5) +
#   theme_classic()

ggplot() +
  geom_density(data = enviro_buffalo_values, aes(x = ndvi, fill = Data), alpha = 0.25) +
  theme_classic()

ggplot() +
  geom_density(data = enviro_buffalo_values, aes(x = slope, fill = Data), alpha = 0.25) +
  theme_classic()


tab <- as.data.frame(prop.table(table(enviro_buffalo_values$canopy, enviro_buffalo_values$Data), 2))
names(tab) <- c("Canopy_cover", "Data", "Freq")

ggplot(data = tab, aes(x = Canopy_cover, y = Freq, fill = Data)) +
  geom_bar(stat = "identity", position = position_dodge(), alpha = 0.75) +
  theme_classic()

```


```{r temporal naive}

years <- c(rep(2019, 7), rep(2018, 5))
months <- c(seq(1, 7, 1), seq(8, 12, 1))
day <- rep(1, 12)
dates <- make_datetime(year = years, month = months, day = day, tz = "Australia/Queensland")

# temp_coefs <- data.frame("date" = dates, 
#                          "ndvi" = harmonic_weighted_mean$NDVI[seq(1, 111, 10)], 
#                          "slope" = harmonic_weighted_mean$Slope[seq(1, 111, 10)], 
#                          "veg_herby" = harmonic_weighted_mean$Herbaceous_vegetation[seq(1, 111, 10)], 
#                          "canopy" = harmonic_weighted_mean$Canopy_cover[seq(1, 111, 10)])

monthly_coefs <- data.frame("date" = dates,
                         "ndvi" = harmonic_coefs_df$NDVI[seq(1, 331, 30)],
                         "ndvi_quad" = harmonic_coefs_df$NDVI_quadratic[seq(1, 331, 30)],
                         "canopy" = harmonic_coefs_df$Canopy_cover[seq(1, 331, 30)],
                         "canopy_quad" = harmonic_coefs_df$Canopy_cover_quadratic[seq(1, 331, 30)],
                         "shape" = harmonic_coefs_df$shape[seq(1, 331, 30)],
                         "scale" = harmonic_coefs_df$scale[seq(1, 331, 30)],
                         "kappa" = harmonic_coefs_df$kappa[seq(1, 331, 30)])

# write_csv(monthly_coefs, "outputs/weighted_mean_monthly_CLR_coefs_wide_yday_with_movement_params_20230424.csv") # weighted mean

monthly_coefs_long <- monthly_coefs %>% pivot_longer(cols = !date)

# sanity check to ensure that the coefficients line up to the correct months
monthly_coefs_long %>% filter(!name %in% c("shape", "scale", "kappa")) %>% 
  ggplot(aes(x = month(date), y = value, colour = factor(name))) +
  geom_line() +
  scale_colour_viridis_d("Covariate") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_x_continuous(breaks = 1:12) +
  theme_classic()

```

# Predictions and cross-validation

Boyce index function

```{r}

ecospat.boyce2 <- function (fit, obs, nclass = 0, window.w = "default", res = 100, 
  PEplot = TRUE, rm.duplicate = TRUE, method = "spearman") 
{
  boycei <- function(interval, obs, fit) {
    pi <- sum(as.numeric(obs >= interval[1] & obs <= interval[2]))/length(obs)
    ei <- sum(as.numeric(fit >= interval[1] & fit <= interval[2]))/length(fit)
    return(round(pi/ei, 10))
  }
  if (inherits(fit, "RasterLayer")) {
    if (is.data.frame(obs) || is.matrix(obs)) {
      obs <- raster::extract(fit, obs)
    }
    fit <- getValues(fit)
    fit <- fit[!is.na(fit)]
    obs <- obs[!is.na(obs)]
  }
  mini <- min(fit, obs)
  maxi <- max(fit, obs)
  if (length(nclass) == 1) {
    if (nclass == 0) {
      if (window.w == "default") {
        window.w <- (max(fit) - min(fit))/10
      }
      vec.mov <- seq(from = mini, to = maxi - window.w, 
        by = (maxi - mini - window.w)/res)
      vec.mov[res + 1] <- vec.mov[res + 1] + 1
      interval <- cbind(vec.mov, vec.mov + window.w)
    }
    else {
      vec.mov <- seq(from = mini, to = maxi, by = (maxi - 
        mini)/nclass)
      interval <- cbind(vec.mov, c(vec.mov[-1], maxi))
    }
  }
  else {
    vec.mov <- c(mini, sort(nclass[!nclass > maxi | nclass < 
      mini]))
    interval <- cbind(vec.mov, c(vec.mov[-1], maxi))
  }
  f <- apply(interval, 1, boycei, obs, fit)
  to.keep <- which(f != "NaN")
  f <- f[to.keep]
  if (length(f) < 2) {
    b <- NA
  }
  else {
    r <- 1:length(f)
    if (rm.duplicate == TRUE) {
      r <- c(1:length(f))[f != c(f[-1], TRUE)]
    }
    b <- cor(f[r], vec.mov[to.keep][r], method = method)
  }
  HS <- apply(interval, 1, sum)/2
  if (length(nclass) == 1 & nclass == 0) {
    HS[length(HS)] <- HS[length(HS)] - 1
  }
  HS <- HS[to.keep]
  if (PEplot == TRUE) {
    plot(HS, f, xlab = "Habitat suitability", ylab = "Predicted/Expected ratio", 
      col = "grey", cex = 0.75)
    points(HS[r], f[r], pch = 19, cex = 0.75)
  }
  return(list(F.ratio = f, cor = round(b, 3), HS = HS))
}

```




```{r}

boyce_naive_list <- vector(mode = "list", length = 12)
naive_pred_stack <- c()

for(i in 1:12) {

resources <- c(ndvi_stack_scaled[[which(time(ndvi_stack_scaled) == monthly_coefs$date[i])]], 
               # slope_scaled,
               # veg_herby, 
               canopy_scaled)

# ndvi
ndvi_lin <- resources[[1]]
ndvi_lin <- ndvi_lin * monthly_coefs$ndvi[[i]]
# plot(ndvi_lin)

ndvi_quad <- resources[[1]]
ndvi_quad <- (ndvi_quad ^ 2) * monthly_coefs$ndvi_quad[[i]] 
# plot(ndvi_quad)

ndvi_pred <- ndvi_lin + ndvi_quad
# plot(ndvi_pred)

# canopy cover 
canopy_lin <- resources[[2]]
canopy_lin <- canopy_lin * monthly_coefs$canopy[[i]]
# plot(canopy_lin)

canopy_quad <- resources[[2]]
canopy_quad <- (canopy_quad ^ 2) * monthly_coefs$canopy_quad[[i]] 
# plot(canopy_quad)

canopy_pred <- canopy_lin + canopy_quad
# plot(canopy_pred)

# combining
naive_pred <- exp(ndvi_pred + canopy_pred) / global(exp(ndvi_pred + canopy_pred), fun = "sum", na.rm = TRUE)[[1]]
plot(naive_pred)
plot(naive_pred)
points(as.numeric(buffalo_CLR_year %>% filter(y == 1 & month == i) %>% dplyr::select(x1) %>% unlist()), 
       as.numeric(buffalo_CLR_year %>% filter(y == 1 & month == i) %>% dplyr::select(y1) %>% unlist()))

naive_pred_stack <- c(naive_pred_stack, naive_pred)


# setting up the buffalo observation data for the relevant month

# buffalo_obs <- buffalo_CLR_year %>% 
#   filter(y == 1 & month == 1 & x1 > xmin & x1 < xmax & y1 > ymin & y1 < ymax) %>% 
#   transmute(x = x1 - xmin, y = y1 - ymin)

buffalo_obs <- buffalo_CLR_year %>% filter(y == 1 & month == i) %>% 
  transmute(x = x1, y = y1)

naive_raster <- raster(naive_pred)
# naive_raster <- raster(naive_pred_cropped)

# plot(naive_pred)
# plot(naive_raster)

# plot(naive_cropped_raster[[1]])

# boyce_naive <- ecospat.boyce2(naive_cropped_raster[[1]], buffalo_obs,
#                               # method = "spearman"
#                               method = "pearson",
#                               # method = "kendall"
#                               )

boyce_naive_list[[i]] <- ecospat.boyce2(naive_raster, buffalo_obs,
                              method = "spearman"
                              # method = "pearson",
                              # method = "kendall"
                              )

boyce_naive_list[[i]]

}

# naive_raster_2harms <- naive_raster
# boyce_naive_list_2harms <- boyce_naive_list

naive_raster_3harms <- naive_raster
boyce_naive_list_3harms <- boyce_naive_list

```


```{r}

for(i in 1:12) {
  plot(naive_pred_stack[[i]])
}

for(i in 1:12) {
  # plot(naive_pred_stack[[i]])
  plot(naive_pred_stack[[i]])
  points(as.numeric(buffalo_CLR_year %>% filter(y == 1 & month == i) %>% dplyr::select(x1) %>% unlist()), 
       as.numeric(buffalo_CLR_year %>% filter(y == 1 & month == i) %>% dplyr::select(y1) %>% unlist()))
}

```


```{r}

spearman_cor <- c()

for(i in 1:12) {
  print(boyce_naive_list[[i]]$cor)
  spearman_cor[i] <- boyce_naive_list[[i]]$cor
}

mean(spearman_cor)
sd(spearman_cor)

```